---
title: "STAT 407 Homework 3"
author: "Charles Hwang"
date: "10/20/2021"
output: pdf_document
---
## Problem 5.6
(a)
```{r Problem 5.6(a)}
rm(list=ls())
current<-c(280,300,290,290,310,285,285,295,290,230,260,220,235,240,225,240,235,230)
glass<-as.factor(rep(1:2,each=9))
phosphor<-as.factor(rep(1:3,times=6))
t<-lm(current~glass*phosphor)
anova(t)
```

We reject both null hypotheses at the $\alpha$ = 0.05 level. There is sufficient evidence (p < 0.000001, p = 0.004364) that both factors influence brightness.

(b)
```{r Problem 5.6(b)}
par(mfrow=c(1,2))
interaction.plot(glass,phosphor,current)
interaction.plot(phosphor,glass,current,ylab="")
# There is interaction between phosphor types 1 and 3. There may be slight interaction
# between phosphor type 2 and the other two phosphor types.
anova(t)["glass:phosphor","Pr(>F)"]
```

We can see the overall interaction is not significant (p = 0.3178005). The two factors do interact, but only at the levels of phosphor types 1 and 3.

(c)
```{r Problem 5.6(c), message=FALSE}
shapiro.test(t$residuals)
bartlett.test(t$residuals~glass)
bartlett.test(t$residuals~phosphor)
library(car)
leveneTest(lm(t))
# None of the null hypotheses were rejected, so the normality and
par(mfrow=c(2,2))  # equal variance assumptions appear to be met.
qqnorm(t$residuals)
qqline(t$residuals)
plot(c(glass),t$residuals,xlab="Glass Type",ylab="Residuals")  
abline(h=0)
plot(c(phosphor),t$residuals,xlab="Phosphor Type",ylab="Residuals")
abline(h=0)
plot(t$fitted.values,t$residuals)
abline(h=0)
# There is a slight variation in the Q-Q plot, but there does not appear to be any pattern
# in the residual plots.
```

## Problem 5.9
```{r Problem 5.9}
force<-c(2.70,2.45,2.60,2.75,2.78,2.49,2.72,2.86,2.83,2.85,2.86,2.94,2.86,2.80,2.87,2.88)
speed<-as.factor(rep(c(125,200),each=8))
feedrate<-as.factor(rep(c(0.015,0.03,0.045,0.06),times=4))
par(mfrow=c(1,2))
interaction.plot(speed,feedrate,force)
interaction.plot(feedrate,speed,force,ylab="")
# We can see that all of the levels have some degree of interaction with each other.
d<-lm(force~speed+feedrate+speed*feedrate)
anova(d)
# We reject both null hypotheses at the alpha = 0.05 level. There is sufficient
# evidence (p = 0.000066, p = 0.002582) that both factors influence thrust force.
shapiro.test(d$residuals)
bartlett.test(d$residuals~speed)
bartlett.test(d$residuals~feedrate)
# None of the null hypotheses were rejected, so the normality and
par(mfrow=c(2,2))  # equal variance assumptions appear to be met.
qqnorm(d$residuals)
qqline(d$residuals)
plot(c(speed),d$residuals,xlab="Drill Speed",ylab="Residuals")  
abline(h=0)
plot(c(feedrate),d$residuals,xlab="Feed Rate",ylab="Residuals")
abline(h=0)
plot(d$fitted.values,d$residuals)
abline(h=0)
# There is a slight variation in the Q-Q plot and a slight pattern in the
# residuals vs. fitted values plot.
TukeyHSD(aov(force~speed+feedrate+speed*feedrate))$speed # Post-hoc analysis
TukeyHSD(aov(force~speed+feedrate+speed*feedrate))$feedrate
# The two drill speeds are clearly different. The 0.03 feed rate appears to be different
# from the 0.015 feed rate and the 0.06 feed rate. The 0.03 feed rate may also be
# different from the 0.045 feed rate, but we narrowly failed to reject the null
# hypothesis at the alpha = 0.05 level (p = 0.050910663).
```

## Problem 5.12
```{r Problem 5.12}
yield<-c(90.4,90.7,90.2,90.2,90.6,90.4,90.1,90.5,89.9,90.3,90.6,90.1,90.5,90.8,90.4,90.7,90.9,90.1)
tempr<-as.factor(rep(c(150,160,170),each=6))
pressure<-as.factor(rep(c(200,215,230),times=6))
TukeyHSD(aov(yield~tempr+pressure+tempr*pressure))$pressure
```

The 215 psig level appears to be different from both the 200 psig level and the 230 psig level.

## Problem 5.13
Model without interaction: $y_{ijk}$ = $\mu$ = $\tau_{i}$ + $\beta_{j}$ + $\epsilon_{ijk}$ (no $(\tau\beta)_{ij}$ term)

i = 1, 2; j = 1, 2, 3; k = 1, 2, 3

```{r Problem 5.13}
density<-c(570,1063,565,565,1080,510,583,1043,590,528,988,526,547,1026,538,521,1004,532)
position<-as.factor(rep(1:2,each=9))
tempc<-as.factor(rep(c(800,825,850),times=6))
a<-lm(density~position+tempc)
anova(a)
# We reject both null hypotheses at the alpha = 0.05 level. There is sufficient
# evidence (p = 0.001254, p < 0.000001) that both factors influence baked density.
shapiro.test(a$residuals)
bartlett.test(a$residuals~position)
bartlett.test(a$residuals~tempc)
# None of the null hypotheses were rejected, but we should exercise slight caution as the
# Shapiro-Wilk normality test is close to the alpha = 0.05 significance level (p = 0.09529).
par(mfrow=c(2,2))
qqnorm(a$residuals)
qqline(a$residuals)
plot(c(position),a$residuals,xlab="Furnace Position",ylab="Residuals")  
abline(h=0)
plot(c(tempc),a$residuals,xlab="Firing Temperature (°C)",ylab="Residuals")
abline(h=0)
plot(a$fitted.values,a$residuals)
abline(h=0)
# There is a slight variation in the Q-Q plot, but not much pattern in the residual plots.
TukeyHSD(aov(density~position+tempc)) # Post-hoc analysis
# We can see that the two furnace positions are different. The 825C level is also clearly
# different from both the 800C level and the 850C level.
```

## "Q"
```{r Q}
par(mfrow=c(1,2))
interaction.plot(position,tempc,density)
interaction.plot(tempc,position,density,ylab="")
anova(lm(density~position+tempc+position*tempc))["position:tempc","Pr(>F)"]
```

Looking at the interaction plots, there does not appear to be any interaction. The assumption of a lack of interaction appears to have been made in good faith and should be considered valid. As such, the model as written appears to be adequate.

## Problem 5.24
(a)
```{r Problem 5.24(a)}
growth<-c(2.29,2.06,1.90,2.47,2.05,1.93,2.48,2.23,1.75,2.12,2.03,2.06,2.65,3.20,3.10,2.68,3.18,3.24,2.06,3.96,3.98,2.38,3.64,3.24,2.24,11.00,9.96,2.71,11.00,10.01,2.81,9.06,9.36,2.08,11.30,10.40)
freq<-as.factor(rep(c(10,1,0.1),each=12))
env<-as.factor(rep(c("Air","H2O","Salt H2O"),times=12))
par(mfrow=c(1,2))
interaction.plot(freq,env,growth)
interaction.plot(env,freq,growth,ylab="")
# We can see that all of the levels have some degree of interaction with each other,
# although the interaction between H2O and salt H2O appears to be slight.
c<-lm(growth~freq*env)
anova(c)
# We reject both null hypotheses at the alpha = 0.05 level. There is sufficient
# evidence (p < 0.000001, p < 0.000001) that both factors influence crack growth rate.
```

(b)
```{r Problem 5.24(b)}
shapiro.test(c$residuals)
bartlett.test(c$residuals~freq)
bartlett.test(c$residuals~env)
# All of the null hypotheses were rejected. The normality and equal variance assumptions
par(mfrow=c(2,2))     # have clearly been violated and this analysis should not be used.
qqnorm(c$residuals)
qqline(c$residuals)
plot(c(freq),c$residuals,xlab="Frequency",ylab="Residuals")  
abline(h=0)
plot(c(env),c$residuals,xlab="Environment",ylab="Residuals")
abline(h=0)
plot(c$fitted.values,c$residuals)
abline(h=0)
# There is a slight variation in the Q-Q plot and there appears to be a megaphone pattern
# in the residuals vs. fitted values plot. This is further evidence that this analysis
# should not be used.
```

(c)
```{r Problem 5.24(c)}
lngrowth<-log(growth)
lnc<-lm(lngrowth~freq+env+freq*env)
anova(lnc)
# We reject both null hypotheses at the alpha = 0.05 level. There is sufficient
# evidence (p < 0.000001, p < 0.000001) that both factors influence crack growth rate.
shapiro.test(lnc$residuals)
bartlett.test(lnc$residuals~freq)
bartlett.test(lnc$residuals~env)
leveneTest(lm(c))
# None of the null hypotheses were rejected, so the normality and
par(mfrow=c(2,2))  # equal variance assumptions appear to be met.
qqnorm(lnc$residuals)
qqline(lnc$residuals)
plot(c(freq),lnc$residuals,xlab="Frequency",ylab="Residuals")  
abline(h=0)
plot(c(env),lnc$residuals,xlab="Environment",ylab="Residuals")
abline(h=0)
plot(lnc$fitted.values,lnc$residuals)
abline(h=0)
# There is a slight variation in the Q-Q plot, but the intensity of the megaphone pattern
# in the residuals vs. fitted values plot appears to have decreased.
TukeyHSD(aov(lngrowth~freq+env+freq*env))$freq # Post-hoc analysis
TukeyHSD(aov(lngrowth~freq+env+freq*env))$env
# We can see that all levels of both factors except H2O and salt H2O for the
# environment factor are clearly different from one another.
```

We saw that the levels of the frequency are powers of 10 (0.1 = $10^{-1}$, 1 = $10^0$, and 10 = $10^1$). A logarithmic transformation of the data appears to have created a more suitable model where the normality and equal variance assumptions are met.

## Extra Credit
Model: $y_{ijk}$ = $\mu$ + $\tau_{i}$ + $\beta_{j}$ + $(\tau\beta)_{ij}$ + $\epsilon_{ijk}$ (i = 1, 2; j = 1, 2, 3; k = 1, 2, 3)

$y_{ijk}$: response variable

$\mu$: overall mean

$\tau_{i}$: level i of factor $\tau$ (method)

$\beta_{j}$: level j of factor $\beta$ (glucose level)

$(\tau\beta)_{ij}$: interaction term

$\epsilon_{ijk}$: error/residual

```{r Extra Credit}
glucose<-c(42.5,138.4,180.9,39.8,132.4,176.8,43.3,144.4,180.5,40.3,132.4,173.6,42.9,142.7,183,41.2,130.3,174.9)
method<-as.factor(rep(1:2,times=3,each=3))
level<-as.factor(rep(1:3,times=6))
par(mfrow=c(1,2))
interaction.plot(method,level,glucose)
interaction.plot(level,method,glucose,ylab="")
# There may be slight interaction between method 1 and method 2.
g<-lm(glucose~method*level)
anova(g)
# We reject both null hypotheses at the alpha = 0.05 level. There is sufficient
# evidence (p = 0.000003, p < 0.000001) that both factors influence concentration of glucose.
shapiro.test(g$residuals)
bartlett.test(g$residuals~method)
bartlett.test(g$residuals~level)
leveneTest(g)
# We should exercise caution here as the null hypothesis for Bartlett's test for homoscedasticity
# by level has been rejected at the alpha = 0.05 level (p = 0.02731).
par(mfrow=c(2,2))
qqnorm(g$residuals)
qqline(g$residuals)
plot(c(method),g$residuals,xlab="Method",ylab="Residuals")  
abline(h=0)
plot(c(level),g$residuals,xlab="Glucose Level",ylab="Residuals")
abline(h=0)
plot(g$fitted.values,g$residuals)
abline(h=0)
# There is a slight variation in the Q-Q plot and a megaphone/football pattern in the
# residuals vs. fitted values plot.
TukeyHSD(aov(glucose~method+level+method*level))$method # Post-hoc analysis
TukeyHSD(aov(glucose~method+level+method*level))$level
# We can see that all levels of both factors are clearly different from one another.
```