---
title: "Homework 1"
author: "Charles Hwang"
date: "9/18/2022"
output: pdf_document
---
## Problem 5

### Problem 5(a)

```{r Problem 5(a)}
rm(list=ls())
NCbirths<-read.csv("/Users/newuser/Desktop/Notes/Graduate/STAT 408 - Applied Regression Analysis/births.csv")
```

### Problem 5(b)

```{r Problem 5(b)}
weights<-NCbirths$weight
```

Given these data are for newborn babies, it appears these weights are listed in ounces.

### Problem 5(c)

```{r Problem 5(c)}
weights_in_pounds<-weights/16
```

### Problem 5(d)

```{r Problem 5(d)}
head(weights_in_pounds,20)
```

### Problem 5(e)

```{r Problem 5(e)}
mean(weights_in_pounds)
```

We can see the mean weight of all babies is `r mean(weights_in_pounds)` pounds.

### Problem 5(f)

```{r Problem 5(f)}
table(NCbirths$Habit)
```

We can see that approximately $\frac{187}{1805+187}=\frac{187}{1992}\approx$ `r as.numeric(100*table(NCbirths$Habit)["Smoker"]/sum(table(NCbirths$Habit)))` percent of mothers in the sample smoke.

### Problem 5(g)

```{r Problem 5(g)}
abs(14-as.numeric(100*table(NCbirths$Habit)["Smoker"]/sum(table(NCbirths$Habit))))
100*(1-as.numeric(100*table(NCbirths$Habit)["Smoker"]/sum(table(NCbirths$Habit)))/14)
```

The percentage I found in problem 5(f) is approximately `r 14-as.numeric(100*table(NCbirths$Habit)["Smoker"]/sum(table(NCbirths$Habit)))` percentage points off from the percentage cited in the CDC report. Alternatively, we can see the percentage from 5(f) is approximately `r 100*(1-as.numeric(100*table(NCbirths$Habit)["Smoker"]/sum(table(NCbirths$Habit)))/14)` percent less than the percentage cited in the CDC report.

## Problem 6

### Problem 6(a)

```{r Problem 6(a)}
flint<-read.csv("/Users/newuser/Desktop/Notes/Graduate/STAT 408 - Applied Regression Analysis/flint.csv")
```

### Problem 6(b)

```{r Problem 6(b)}
mean(flint$Pb>=15)
```

We can see that approximately `r 100*mean(flint$Pb>=15)` percent of the locations tested were found to have dangerous lead levels.

### Problem 6(c)

```{r Problem 6(c)}
mean(flint[flint$Region=="North","Cu"])
```

### Problem 6(d)

```{r Problem 6(d)}
mean(flint[flint$Region=="North" & flint$Pb>=15,"Cu"])
```

### Problem 6(e)

```{r Problem 6(e)}
apply(flint[,c("Pb","Cu")],2,mean)
```

### Problem 6(f)

```{r Problem 6(f)}
boxplot(flint$Pb,horizontal=TRUE,main="Lead (Pb) Levels at Various Locations in Flint, Michigan",xlab="Lead Level in PPB (parts per billion)")
```

### Problem 6(g)

```{r Problem 6(g)}
mean(flint$Pb>0)
```

No, the mean does not seem to be a good measure of center for these data. A more useful statistic for these data could be the proportion of locations at which *any* level of lead was measured (`r mean(flint$Pb>0)`), or the proportion found in problem 6(b).

## Problem 7

### Problem 7(a)

```{r Problem 7(a)}
set.seed(2022)
hist(NCbirths$weight,main="Problem 7(a) - Histogram of Newborn Baby Weights",xlab="Weight (ounces)")
```

The weight variable appears to be approximately normal, but probably does not *exactly* follow a normal distribution. We can see the shape is symmetric and approximately normal, but the long left tail is a notable sign of non-normality.

### Problem 7(b)

```{r Problem 7(b)}
set.seed(2022)
mean(sample(NCbirths$weight,10))
```

### Problem 7(c)

```{r Problem 7(c)}
set.seed(2022)
m<-m30<-m100<-rep(NULL,1000)
for (i in 1:1000){m[i]<-mean(sample(NCbirths$weight,10))}
hist(m,main="Problem 7(c) - Histogram of Means of Newborn Baby Weights",xlab="Weight (ounces)")
```

Yes, this distribution is close to normal.

### Problem 7(d)

```{r Problem 7(d)}
set.seed(2022)
for (i in 1:1000){m30[i]<-mean(sample(NCbirths$weight,10))}
for (i in 1:1000){m100[i]<-mean(sample(NCbirths$weight,10))}
par(mfrow=c(1,2))
hist(m30,main="Newborn Baby Weights (n=30)",xlab="Weight (ounces)")
hist(m100,main="Newborn Baby Weights (n=100)",xlab="Weight (ounces)",ylab="")
```

Yes, these two distributions are both close to normal. We can see the shape is symmetric and consistent with distributions that are approximately normal.