---
title: "STAT 488 Project 1: Current Employees of the City of Chicago"
author: "Charles Hwang"
date: "4/3/2022"
output: pdf_document
---
The dataset I chose for this project is the "Current Employee Names, Salaries, and Position Titles" dataset from the Chicago Data Portal. I chose this dataset because I am more confident that this data is more accurate than an average dataset, due to it being public data and thus potentially being required to be reported (e.g., through FOIA requests). This dataset is updated every six months (with the last update on January 21, 2022) and contains eight variables: the (1) name, (2) job title, (3) department, (4-5) statuses (full- vs. part-time and salary vs. hourly), (6-7) hours per week and hourly wage (if hourly), and (8) annual salary (if salaried) of every current employee working for the City of Chicago.

I first shortened the variable names, coded the status variables (FTPT and S) as factors, and estimated annual salary for hourly workers by multiplying 50 by hours per week (HpW) by hourly wage (DpH), where 50 is the number of workweeks in a calendar year when excluding city holidays[^1]. A summary of the variables is in **Figure 1**. There are 249 employee names that appear more than once (**Figure 2**), which could be from an employee holding multiple positions or multiple employees having the same first and last name. There are up to 1078 unique[^2] jobs (**Figure 3**), with "Police Officer" being the most common (8876) and a strong statistical outlier. There are 36 departments (**Figure 4**) which match those listed on the city's website (<https://www.chicago.gov/city/en/depts.html>).

I created a two-way table to compare statuses (**Figure 5**) and as expected, there are no part-time salaried employees. I also created tables for hours per week (**Figure 6**) and saw that there are only four levels: 40, 35, 20, and 10. A large majority of hourly employees work 40 hours per week. I also created boxplots to visualize annual salaries (**Figure 7**) and we can see that salaried employees generally make more than hourly employees per year and full-time employees generally make more than part-time employees per year, as expected. The boxplots for salaried employees and full-time employees are very similar because most full-time employees are salaried (as seen in **Figure 6**).

I created tables (**Figure 8**) for hourly employees and interestingly, there are 238 employees listed as being paid below minimum wage (15 DpH[^3] in Chicago). There is one employee each with hourly wages of 12 and 14 DpH, which could be from when the minimum wage was lower or an error with data entry. There are 236 employees listed as being paid 3 DpH, which could be a placeholder for interns, unpaid positions with stipends, volunteers, etc. The boxplots (**Figure 9**) comparing hourly wages by status and workload show all 236 are working 20 hours per week. We can also see that full-time employees are generally paid more than part-time employees, as expected.

With salary (Sal) as the response variable, I ran one GLM with the two factor variables (FTPT and S) as explanatory variables and another with the two quantitative variables (HpW and DpH) as explanatory variables (**Figure 10**). The intercept and both variables were statistically significant in both models. We can see that part-time employees made approximately 69343.47 per year less than full-time employees and salaried employees made approximately 13678.66 per year more than hourly employees (both while fixing the other variable). In the other model, salary increased by approximately 1426.20 per year for every additional hour worked per week when fixing hourly wage.

Q-Q plots (**Figure 11**) showed that none of the three quantitative variables are normally distributed, which is expected with the right skew seen in the boxplots. Subsequent `sqrt` and `log` transformations (not shown) on affected variables are insufficient for normalizing the data. After printing a scatterplot matrix and histograms (**Figure 12**) for the three variables using the `mvn()` function, I conducted a Box-Cox transformation and produced a chi-square Q-Q plot which showed the data were still not normal, but perhaps closer to being normal than before with there being only two visual outliers with unduly large Mahalanobis distances.

I also calculated the means and variance matrix of hours per week (HpW) and hourly wages (DpH) for hourly employees (excluding Sal because it's a proxy variable). Even though the normality assumption is clearly violated, I decided to compute Hotelling's $T^2$ (**Figure 13**) to see what the results would have been and to conduct some multivariate analysis. Since there are so many values in the data (6814), any deviation from the mean in any variable leads to very high values for $T^2$ and strong evidence against nearly any nontrivial matrix $\mu_0=\begin{bmatrix}x_{HpW}\\x_{DpH}\end{bmatrix}$ used for $H_0$. The corresponding $F$-distribution[^4] is $\frac{(n-1)p}{n-p}F_{p,n-p}\approx2.000294F_{2,6810}$, and the critical value at $\alpha=0.05$ is only 5.99498. Because of this, I decided to instead see what values for each individual variable would lead to rejection. We can see that an 11 minute 44 second (or greater) increase in mean weekly workload *or* a 31-cent (or greater) increase in mean hourly wage is significant at the $\alpha=0.05$ level. Evidently, if we were to calculate differences using both variables, even less deviation from the mean would be needed to be statistically significant.

I attempted to do PCA on the same data (**Figure 14**), but since there were only two variables it produced limited results (something I discuss in the conclusion). It was able to produce variance proportions and graphing parameters, but the scree plot was soundly underwhelming. So in an 11th-hour bid to do some meaningful multivariate analysis, I leaned on my knowledge from Predictive Analytics and grew a decision tree (**Figure 15**) and random forest (**Figure 16**) to predict salary (Sal) on the data for hourly employees. It was hard to interpret anything useful from the decision tree due to the Name variable confounding the results, but we can see the error accounted for from each variable in the complexity parameter graph and the table of predictions. We can also see the cross-validation error is relatively low at 0.2615813.

The plot of log-transformed error shows the random forest was properly fit after around 200 trees. We can also see from the variable importance plot that and hourly wages (DpH) is the most important variable, as expected. This is followed by workload (HpW), status (S), and department (Dept) which makes sense intuitively. Different departments in the City of Chicago have different pay grades due to public pensions, labor unions, staffing/hiring levels, etc.

Job, Name, and status (which is trivial because this subset is entirely hourly employees) round out the explanatory variables in the random forest. The importance of the Job variable has a similar reasoning as department (Dept), but in future analysis I would be interested to see if the Name variable has more importance or influence. Specifically, seeing if certain "sounding" names imply a certain race or gender which may lead to inherent bias in hiring practices (and thus salary). There was no Race or Sex variable so this would likely require a deep-dive into the Name variable or using some other prediction/machine learning function to subset it. This is one of the few things I would explore in future analysis if given the time.

Another thing I would do if I had additional time would be to clean the Job variable. It was mentioned in footnote 2 that there may be duplicates and we saw already (**Figure 3**) that there are two mentions each of police officers and firefighter (EMT) in different roles just in the ten most common positions. Reclassifying entries listed in positions with only a few employees to more common positions or categorizing positions by profession/occupation, type of work, etc. could also be helpful. This seems to be easier than the deep-dive into the Name variable but still a challenge with 1078 levels.

In hindsight, better review and selection of data and better understanding of the project would have produced more thorough and robust analysis. At the time of choosing this dataset, it did not occur to me that there were only three quantitative variables and that they would all have missing values (due to employees either being salaried or hourly). Given the history of Chicago public affairs, it seems this dataset may be more used for looking up details on specific employees (i.e., politicians, lobbyists, etc.) rather than any statistial analysis. The normality assumption was also badly violated (**Figures 11-12**) as the quantitative variables were all right-skew, so some analyses may not be valid. It seemed sufficient after reviewing the sample size and variables and seeing it was the most viewed and downloaded dataset in the Chicago Data Portal, but I did not do the best job of vetting the dataset for the analysis needed. However, I believe the project was a success overall given the short timeline in applying multivariate analyses to real-world data.

[^1]: According to the city's website (<https://www.chicago.gov/city/en/narr/misc/city-holidays.html>), there are 12 holidays on which offices are closed. Since there are approximately 52 weeks in a year, subtracting 12 days results in *roughly* 50 weeks.

[^2]: I say "up to" because there may be jobs that are entered under different names, causing additional levels to be created.

[^3]: Unfortunately, RMarkdown does not allow me to (easily) type dollar signs as it interprets them as the start of a LaTeX equation, but the units for the hourly wage (DpH) variable are all in United States dollars per hour.

[^4]: We can see that (*n*-1)*p*/(*n*-*p*) converges to *p* for large *n*, as is the case here.

## Appendix

### Figure 1

```{r Figure 1}
rm(list=ls())
cdp<-read.csv("/Users/newuser/Desktop/Notes/Graduate/STAT 488 - Multivariate Statistical Analysis/Current_Employee_Names__Salaries__and_Position_Titles.csv")
names(cdp)<-c("Name","Job","Dept","FTPT","S","HpW","Sal","DpH")
cdp$Dept<-as.factor(cdp$Dept) # 36 departments
cdp$FTPT<-as.factor(cdp$FTPT)
cdp$S<-as.factor(cdp$S)
cdp$Sal[is.na(cdp$Sal)==TRUE]<-50*cdp$DpH[is.na(cdp$DpH)==FALSE]*cdp$HpW[is.na(cdp$HpW)==FALSE]
summary(cdp)
```

### Figure 2

```{r Figure 2}
head(sort(table(cdp$Name),decreasing=TRUE),32)
```

### Figure 3

```{r Figure 3}
head(sort(table(cdp$Job),decreasing=TRUE),0.01*length(table(cdp$Job)))
```

### Figure 4

```{r Figure 4}
sort(table(cdp$Dept),decreasing=TRUE)
```

### Figure 5

```{r Figure 5}
xtabs(~FTPT+S,data=cdp)[,c("Salary","Hourly")]
```

### Figure 6

```{r Figure 6}
table(cdp$HpW)[sort(names(table(cdp$HpW)),decreasing=TRUE)] # HpW (overall)
table(cdp$HpW[cdp$FTPT=="F"])[sort(names(table(cdp$HpW)),decreasing=TRUE)] # HpW (FT)
table(cdp$HpW[cdp$FTPT=="P"])[sort(names(table(cdp$HpW)),decreasing=TRUE)] # HpW (PT)
```

### Figure 7

```{r Figure 7}
par(mfrow=c(1,3))
boxplot(cdp$Sal,ylab="Salary (in US$ per year)",main="(Prorated) Annual Salaries")
boxplot(cdp$Sal[cdp$S=="Salary"],cdp$Sal[cdp$S=="Hourly"],names=c("Salary","Hourly"),main="(Prorated) Salaries by Status")
boxplot(cdp$Sal[cdp$FTPT=="F"],cdp$Sal[cdp$FTPT=="P"],names=c("Full-Time","Part-Time"),main="(Prorated) Salaries by Status")
```

### Figure 8

```{r Figure 8}
head(sort(table(cdp$DpH),decreasing=TRUE))
head(table(cdp$DpH)[as.character(sort(as.numeric(names(table(cdp$DpH)))))])
```

### Figure 9

```{r Figure 9}
par(mfrow=c(1,3))
boxplot(cdp$DpH,ylab="Wages (in US$ per hour)",main="Hourly Wages")
abline(h=15,lty=2) # Minimum wage is $15.00 per hour
boxplot(cdp$DpH[cdp$FTPT=="F"],cdp$DpH[cdp$FTPT=="P"],names=c("Full-Time","Part-Time"),xlab="Status",main="Hourly Wages by Status")
abline(h=15,lty=2)
boxplot(subset(cdp$DpH,cdp$HpW==40),subset(cdp$DpH,cdp$HpW==35),subset(cdp$DpH,cdp$HpW==20),subset(cdp$DpH,cdp$HpW==10),names=sort(names(table(cdp$HpW)),decreasing=TRUE),xlab="Hours per Week",main="Hourly Wages by Workload")
abline(h=15,lty=2)
```

### Figure 10

```{r Figure 10}
summary(glm(Sal~FTPT+S,data=cdp)) # Factor variables
summary(glm(Sal~HpW+DpH,data=cdp)) # Quantitative variables
```

### Figure 11

```{r Figure 11}
par(mfrow=c(1,3))
qqnorm(cdp$HpW)
qqline(cdp$HpW)
qqnorm(cdp$Sal,ylab="")
qqline(cdp$Sal)
qqnorm(cdp$DpH,ylab="")
qqline(cdp$DpH)
```

### Figure 12

```{r Figure 12}
cdpn<-cdp[,c("HpW","Sal","DpH")]
library(MVN)
mvn(cdpn,univariatePlot="scatter")$Descriptives
mvn(cdpn,univariatePlot="histogram")$Descriptives
mvn(cdpn,multivariatePlot="qq",bc=TRUE)$BoxCoxPowerTransformation
```

### Figure 13

```{r Figure 13}
x<-colMeans(cdpn[is.na(cdpn$HpW)==FALSE,c("HpW","DpH")])
x
cor(cdpn[is.na(cdpn$HpW)==FALSE,c("HpW","DpH")])
vm<-cov(cdpn[is.na(cdpn$HpW)==FALSE,c("HpW","DpH")])
vm
n<-dim(cdpn[is.na(cdpn$HpW)==FALSE,c("HpW","DpH")])[1]
p<-length(cdpn[,c("HpW","DpH")])
(n-1)*p/(n-p)*qf(1-0.05,p,n-p)
n*t(x-round(x))%*%solve(vm)%*%(x-round(x))
n*t(matrix(c(11/60+44/60/60,0)))%*%solve(vm)%*%(matrix(c(11/60+44/60/60,0)))
n*t(matrix(c(0,0.31)))%*%solve(vm)%*%(matrix(c(0,0.31)))
```

### Figure 14

```{r Figure 14}
library(purrr)
prcomp(cdpn[is.na(cdpn$HpW)==FALSE,c("HpW","DpH")],scale=TRUE)
summary(prcomp(cdpn[is.na(cdpn$HpW)==FALSE,c("HpW","DpH")],scale=TRUE))
as.data.frame(map(prcomp(cdpn[is.na(cdpn$HpW)==FALSE,c("HpW","DpH")],scale=TRUE),sd))
plot(predict(prcomp(cdpn[is.na(cdpn$HpW)==FALSE,c("HpW","DpH")],scale=TRUE)))
screeplot(prcomp(cdpn[is.na(cdpn$HpW)==FALSE,c("HpW","DpH")],scale=TRUE),type="lines")
```

### Figure 15

```{r Figure 15, warning=FALSE}
library(rpart)
set.seed(3422)
tree<-rpart(Sal~.,data=cdp)
ptree<-prune(tree,cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
printcp(ptree)
plotcp(ptree)
plot(ptree,uniform=TRUE,main="Pruned Tree")
text(ptree,cex=0.6)
table(predict(ptree))
min(tree$cptable[,"xerror"]) # Cross-validation error
```

### Figure 16

```{r Figure 16, message=FALSE}
library(randomForest)
set.seed(3422)
rf<-randomForest(Sal~.,data=cdp[!is.na(cdp$HpW),],ntree=1000,importance=TRUE)
rf
plot(log(rf$mse),type="l",main="Random Forest Error")
varImpPlot(rf)
```

## Works Cited

| "Current Employee Names, Salaries, and Position Titles." *Chicago Data Portal*, 27 Sept. 2011,
|         <https://data.cityofchicago.org/Administration-Finance/Current-Employee-Names-Salaries-and-Position-Title/xzkq-xp2w>.
|         Accessed 3 April 2022.