---
title: "STAT 488: Multivariate Statistical Analysis — Homework 1"
author: "Charles Hwang"
date: "2/3/2022"
output: pdf_document
---
## Problem 1.1

```{r Problem 1.1, tidy=TRUE, tidy.opts=list(width.cutoff=80)}
x1<-matrix(c(3,4,2,6,8,2,5))
x2<-matrix(c(5,5.5,4,7,10,5,7.5))
r1<-data.frame(mean(x1),mean(x2),sum((x1-mean(x1))^2)/length(x1),sum((x2-mean(x2))^2)/length(x2),sum((x1-mean(x1))*(x2-mean(x2)))/length(x1))
names(r1)<-c("x1","x2","s11","s22","s12")
r1
```

## Problem 1.4

```{r Problem 1.4}
x1<-matrix(c(108.28,152.36,95.04,65.45,62.97,263.99,265.19,285.06,92.01,165.68))
x2<-matrix(c(17.05,16.59,10.91,14.14,9.52,25.33,18.54,15.73,8.10,11.13))
x3<-matrix(c(1484.10,750.33,766.42,1110.46,1031.29,195.26,193.83,191.11,1175.16,211.15))
plot(x1,x2,main="Problem 1.4(a)")          # Problem 1.4(a)
# There appears to be positive correlation between x1 and x2.
s11<-sqrt(sum((x1-mean(x1))^2)/length(x1)) # Problem 1.4(b)
s22<-sqrt(sum((x2-mean(x2))^2)/length(x2))
s12<-s21<-sum((x1-mean(x1))*(x2-mean(x2)))/length(x1)
r4<-data.frame(mean(x1),mean(x2),s11,s22,s12,cor(x1,x2))
names(r4)<-c("x1","x2","s11","s22","s12","cor(x1,x2)")
r4
```

There is moderate correlation (*r* = `r cor(x1,x2)`) between $x_1$ and $x_2$.

## Problem 1.5

```{r Problem 1.5}
par(mfrow=c(1,2))                                     # Problem 1.5(a)
plot(x2,x3)
plot(x1,x3)
# There appears to be weak negative correlation between x2 and x3
# and a moderately strong negative correlation between x1 and x3.
s13<-s31<-sum((x1-mean(x1))*(x3-mean(x3)))/length(x1) # Problem 1.5(b)
s23<-s32<-sum((x2-mean(x2))*(x3-mean(x3)))/length(x2)
s33<-sqrt(sum((x3-mean(x3))^2)/length(x3))
x<-matrix(c(mean(x1),mean(x2),mean(x3)))
x
Sn<-matrix(c(s11,s12,s13,s21,s22,s23,s31,s32,s33),nrow=3,ncol=3)
Sn
```
```{r, tidy=TRUE, tidy.opts=list(width.cutoff=80)}
R<-matrix(c(cor(x1,x1),cor(x1,x2),cor(x1,x3),cor(x2,x1),cor(x2,x2),cor(x2,x3),cor(x3,x1),cor(x3,x2),cor(x3,x3)),nrow=3,ncol=3)
R
```

## Problem 1.9

```{r Problem 1.9}
x1<-c(-6,-3,-2,1,2,5,6,8)                                              # Problem 1.9(a)
x2<-c(-2,-3,1,-1,2,1,5,3)
plot(x1,x2,main="Problem 1.9(a)")
s11<-sum((x1-mean(x1))^2)/length(x1)
s22<-sum((x2-mean(x2))^2)/length(x2)
s12<-sum((x1-mean(x1))*(x2-mean(x2)))/length(x1)
r9a<-data.frame(s11,s22,s12)
r9a
x1t<-x1*.899+x2*.438
x2t<-x2*.899-x1*.438
r9b<-data.frame(t(data.frame(x1t,x2t)))                                # Problem 1.9(b)
names(r9b)<-NULL
r9b
s11t<-sum((x1t-mean(x1t))^2)/length(x1t)                               # Problem 1.9(c)
s22t<-sum((x2t-mean(x2t))^2)/length(x2t)
r9c<-data.frame(s11t,s22t)
r9c
sqrt((4*.899+(-2)*.438)^2/s11t+((-2)*.899-4*.438)^2/s22t)              # Problem 1.9(d)
```
```{r, tidy=TRUE, tidy.opts=list(width.cutoff=74)}
a11<-.899^2/(.899^2*s11+2*.438*.899*s12+.438^2*s22)+.438^2/(.899^2*s22-2*.438*.899*s12+.438^2*s11) # Problem 1.9(e)
a22<-.438^2/(.899^2*s11+2*.438*.899*s12+.438^2*s22)+.899^2/(.899^2*s22-2*.438*.899*s12+.438^2*s11)
a12<-.438*.899/(.899^2*s11+2*.438*.899*s12+.438^2*s22)-.438*.899/(.899^2*s22-2*.438*.899*s12+.438^2*s11)
sqrt(a11*(4)^2+2*a12*(4)*(-2)+a22*(-2)^2)
```

We can see this is the same distance as the one using the equation in Problem 1.9(d).

## Problem 1.12

```{r Problem 1.12}
max(abs(-3),abs(4))                                                           # Problem 1.12(a)
plot(NA,xlim=c(-3,3),ylim=c(-2,2),xlab="x1",ylab="x2",main="Problem 1.12(b)") # Problem 1.12(b)
rect(-1,-1,1,1,border="black",col="gray90")
points(0,0)
```

Observing the equation for distance, *d*(*O*,*P*) = max($|x_1|$,$|x_2|$), we can clearly see "the locus of points" is simply a square bounded by *x* = -1, 1 and *y* = -1, 1.

## Problem 1.15

```{r Problem 1.15, tidy=TRUE, tidy.opts=list(width.cutoff=78)}
x1<-c(.889,2.813,1.454,.294,2.727,3.937,2.786,5.231,1.15,6.5,.8,4.6,3.5,3.444,4.071,3.692,5.167,.5,2.385,2.1,5,4.571,2.733,4.235,0,.75,3.077,1.6,6.273,2.625,1.25,2.437,4.454,.133,.222,2.467,4,5.385,.773,3.786,1.923,1,5.8,6.062,3.706,2.444,6.111,2.533,2.167,2.375,1.875,1.75,7.333,5.25,5.182,1.875,5.4,1.154,6.375,9.454,1,1.444,1.8,2.818,10.461,4.143,1.227,5.667,4.111,4.444,3.714,7.4,3.182,5.2,2.333,3.333,5.25,7.714,3.846,2.444,5.333,1.556,3.182,6.222,7.231,3.857,3.778,6,2.333,7.571,3.667,3.6,3.364,4.1,.125,6.231,3,.889)
x2<-c(1.389,1.437,1.091,.941,2.545,1.25,1.714,2.692,1.1,2.562,1,2,1.286,2.556,1,1,3,1,1.923,1.3,3.25,1.214,1.133,2.294,1,1.125,1.462,1.2,3.636,1,1,2.062,1.727,1,1.222,2.667,1,3.154,1,2,1.615,1.333,2.6,1,1.235,2.333,2.222,1.067,1,1.062,1.312,1.333,1.333,1.375,2,2,2,1,2.25,2.727,1,1.111,1.1,2,2.154,1.929,1.182,3,2.556,1.667,3.857,3.7,2.455,2.6,1.667,1.917,2.75,4,2.615,1.111,1.917,1.778,1.545,2.444,1,1.071,1.944,1.4,3.583,2.143,2,2.933,1.273,1.9,1.062,2.769,1.455,1)
x3<-c(1.555,.999,2.364,1.059,2.819,1.937,2.357,1.077,.95,1.749,2.2,3,2.714,2.388,1,2.538,1,1,2.539,1.3,3.125,3.286,2.6,2.706,1.941,3,2.384,2.95,1.182,2.438,2,1.687,2.637,1,1.445,2.2,4,2.384,2.273,1.571,1.693,1.834,3,1.562,1.53,1.223,2.889,1.6,2.167,2.375,2.188,1.167,1.459,2.812,2.727,2.25,1.2,1.923,2.5,3.818,1.917,2,3.1,1.955,2.769,2.642,1.091,1.667,2.222,2.222,2.643,3.1,1.636,.8,.666,2.083,2.5,3.071,3,1,3,3.444,1.91,3.689,3.154,3,1.612,2.067,2.334,3.143,2.111,2.067,1.81,2.8,1.437,1.462,2.09,1)
x4<-c(2.222,2.312,2.455,2,2.727,2.937,2.071,1.846,2,2.562,2.267,2.5,1.286,2.389,2.357,2.154,2.667,2,2.154,1.8,2.375,2.571,1.933,2.176,2,1.875,2,2,2.545,1.937,2,1.875,2.636,2,2,1.933,2.167,2.846,1.909,1.786,2,2,2.8,2.375,2.118,2.444,2.889,2,2,2,2.125,1.75,1.958,2.125,2.818,2.813,1.8,1.846,2.125,2.455,1.833,2.111,2.2,2.045,2,2.429,2.227,2.667,2.778,2,2.286,2.5,2.273,1.8,1.667,1.917,2,2.929,2.692,2.111,2.25,2.667,2.273,2.444,2.308,2.071,1.611,2.267,2.333,2.571,2.778,2.2,2,2,1.875,2.385,2.273,2)
x5<-c(1.945,2.312,2.909,1,4.091,3.749,2,2.539,1,2.499,2.466,3.4,1.252,3,1.572,2.615,3.666,1,2.461,2.6,3.375,3.572,1.667,1.883,2,2,1.846,2.75,3.364,2.062,3,1.375,3.546,1,1,1.8,2.5,2.539,2.091,1.285,1.846,1.917,4.2,1.75,2.294,1.776,3.555,1.333,2.5,2.125,2.062,1,1.542,2.563,4,2.437,1.4,2.462,3,3.272,2.167,2,2.6,2.546,2.923,3.142,3.182,5,3.778,2.444,3.285,4.2,3,2,2.166,3,4,4.428,3.693,1.667,1.917,3.333,3,3.445,4.384,2.286,1.945,2.866,2.667,3.929,4,2.867,2.273,2.6,1.563,4,3.272,1)
x6<-c(1,2,3,1,0,1,2,1,1,1,2,1,3,1,1,1,0,0,1,1,0,1,1,1,0,3,2,1,0,2,1,1,1,0,1,3,0,1,0,3,1,1,1,0,1,3,2,1,1,3,2,1,3,3,2,2,2,1,1,3,1,1,1,2,0,3,1,1,1,0,0,1,1,0,0,1,0,3,2,2,1,1,1,1,2,1,1,2,2,1,3,0,0,2,0,2,2,2)
plot(x2,x3,main="Problem 1.15(a)")
x<-matrix(c(mean(x1),mean(x2),mean(x3),mean(x4),mean(x5),mean(x6))) # Problem 1.15b
x
Sn<-matrix(c(sum((x1-mean(x1))^2)/length(x1),sum((x1-mean(x1))*(x2-mean(x2)))/length(x1),sum((x1-mean(x1))*(x3-mean(x3)))/length(x1),sum((x1-mean(x1))*(x4-mean(x4)))/length(x1),sum((x1-mean(x1))*(x5-mean(x5)))/length(x1),sum((x1-mean(x1))*(x6-mean(x6)))/length(x1),sum((x1-mean(x1))*(x2-mean(x2)))/length(x1),sum((x2-mean(x2))^2)/length(x2),sum((x2-mean(x2))*(x3-mean(x3)))/length(x2),sum((x2-mean(x2))*(x4-mean(x4)))/length(x2),sum((x2-mean(x2))*(x5-mean(x5)))/length(x2),sum((x2-mean(x2))*(x6-mean(x6)))/length(x2),sum((x1-mean(x1))*(x3-mean(x3)))/length(x1),sum((x2-mean(x2))*(x3-mean(x3)))/length(x2),sum((x3-mean(x3))^2)/length(x3),sum((x3-mean(x3))*(x4-mean(x4)))/length(x3),sum((x3-mean(x3))*(x5-mean(x5)))/length(x3),sum((x3-mean(x3))*(x6-mean(x6)))/length(x3),sum((x1-mean(x1))*(x4-mean(x4)))/length(x1),sum((x2-mean(x2))*(x4-mean(x4)))/length(x2),sum((x3-mean(x3))*(x4-mean(x4)))/length(x3),sum((x4-mean(x4))^2)/length(x4),sum((x4-mean(x4))*(x5-mean(x5)))/length(x4),sum((x4-mean(x4))*(x6-mean(x6)))/length(x4),sum((x1-mean(x1))*(x5-mean(x5)))/length(x1),sum((x2-mean(x2))*(x5-mean(x5)))/length(x2),sum((x3-mean(x3))*(x5-mean(x5)))/length(x3),sum((x4-mean(x4))*(x5-mean(x5)))/length(x4),sum((x5-mean(x5))^2)/length(x5),sum((x5-mean(x5))*(x6-mean(x6)))/length(x5),sum((x1-mean(x1))*(x6-mean(x6)))/length(x1),sum((x2-mean(x2))*(x6-mean(x6)))/length(x2),sum((x3-mean(x3))*(x6-mean(x6)))/length(x3),sum((x4-mean(x4))*(x6-mean(x6)))/length(x4),sum((x5-mean(x5))*(x6-mean(x6)))/length(x5),sum((x6-mean(x6))^2)/length(x6)),nrow=6,ncol=6)
Sn
R<-matrix(c(cor(x1,x1),cor(x1,x2),cor(x1,x3),cor(x1,x4),cor(x1,x5),cor(x1,x6),cor(x2,x1),cor(x2,x2),cor(x2,x3),cor(x2,x4),cor(x2,x5),cor(x2,x6),cor(x3,x1),cor(x3,x2),cor(x3,x3),cor(x3,x4),cor(x3,x5),cor(x3,x6),cor(x4,x1),cor(x4,x2),cor(x4,x3),cor(x4,x4),cor(x4,x5),cor(x4,x6),cor(x5,x1),cor(x5,x2),cor(x5,x3),cor(x5,x4),cor(x5,x5),cor(x5,x6),cor(x6,x1),cor(x6,x2),cor(x6,x3),cor(x6,x4),cor(x6,x5),cor(x6,x6)),nrow=6,ncol=6)
R
```

We can see the correlations between most but not all pairs of variables are positive and vary from no to moderate strength. The correlation between amount of food eaten and appetite (*r* = `r cor(x4,x5)`) is the strongest.

## Problem 1.17

```{r Problem 1.17, tidy=TRUE, tidy.opts=list(width.cutoff=71)}
m1<-c(11.57,11.12,11.15,11.14,11.46,11.17,10.98,11.65,10.79,11.31,12.52,11.72,11.09,11.42,11.63,11.13,10.73,10.81,11.1,10.83,11.92,11.41,11.56,11.38,11.43,11.45,11.14,11.36,11.62,11.49,11.8,11.76,11.5,11.72,11.09,11.66,11.08,11.32,11.41,11.96,11.28,10.93,11.3,11.3,10.77,12.38,12.13,11.06,11.16,11.34,11.22,11.33,11.25,10.49)
m2<-c(22.94,22.23,22.7,22.48,23.05,22.6,22.62,23.84,22.01,22.92,25.91,23.92,21.97,23.36,23.91,22.39,21.99,21.71,22.1,22.67,24.5,23.06,23.86,22.82,23.02,23.15,22.6,23.33,23.37,23.8,25.1,23.96,23.37,23.83,23.13,23.69,22.81,23.13,23.31,24.68,23.35,22.13,22.88,22.35,21.87,25.45,24.54,22.38,22.82,22.88,22.56,23.3,22.71,21.34)
m4<-c(52.5,48.63,50.62,51.45,53.3,50.62,49.91,53.68,49.81,49.64,61.65,52.57,47.99,52.92,53.02,50.14,48.25,47.6,49.43,50.56,55.64,51.5,55.08,51.05,51.07,52.06,51.31,51.93,51.56,53.67,56.23,56.07,52.56,54.62,48.89,52.96,51.35,51.6,52.45,55.18,54.75,49.28,51.92,49.88,49.11,56.32,55.08,49.67,51.69,51.32,52.74,52.6,53.15,48.83)
m8<-c(2.05,1.98,1.94,1.97,2.07,1.97,1.97,2,1.93,2.04,2.28,2.1,1.89,2.02,2.09,2.01,1.94,1.92,1.94,2,2.15,1.99,2.1,2,2.01,2.07,1.96,2.01,1.97,2.09,1.97,2.07,2.12,2.06,2.02,2.03,1.93,1.97,2.03,2.24,2.12,1.95,1.98,1.92,1.91,2.29,2.12,1.96,1.99,1.98,2.08,2.06,2.01,1.94)
m15<-c(4.25,4.02,4.05,4.08,4.29,4.17,4,4.22,3.84,4.34,4.82,4.52,4.03,4.12,4.54,4.1,4.03,3.96,3.97,4.09,4.48,4.02,4.36,4.1,3.98,4.24,3.98,4.16,3.96,4.24,4.25,4.35,4.39,4.33,4.19,4.2,4.06,4.1,4.01,4.62,4.41,3.99,3.96,3.9,3.87,5.42,4.52,4.01,4.09,3.97,4.38,4.38,3.92,3.95)
m30<-c(9.19,8.63,8.78,8.82,9.81,9.04,8.54,9.26,8.1,9.37,11.1,9.84,8.87,8.71,9.89,8.69,8.64,8.51,8.37,8.96,9.71,8.55,9.5,9.11,8.36,9.33,8.59,8.74,8.39,9.01,8.96,9.21,9.31,9.24,8.89,9.08,8.57,8.76,8.53,10.21,9.81,8.53,8.5,8.36,8.38,13.12,9.94,8.48,8.81,8.6,9.63,10.07,8.53,8.43)
mt<-c(150.32,143.51,154.35,143.05,174.18,147.41,148.36,152.23,139.39,155.19,212.33,164.33,145.19,149.34,166.46,148.00,148.27,141.45,135.25,153.40,171.33,148.50,154.29,158.10,142.23,156.36,143.47,139.41,138.47,146.12,145.31,149.23,169.28,167.09,144.06,158.42,143.43,146.46,141.06,221.14,165.48,144.18,143.29,142.50,141.31,191.58,154.41,146.51,150.39,145.51,159.53,162.39,151.43,141.16)
x<-matrix(c(mean(m1),mean(m2),mean(m4),mean(m8),mean(m15),mean(m30),mean(mt)))
x
Sn<-round(matrix(c(sum((m1-mean(m1))^2)/length(m1),sum((m1-mean(m1))*(m2-mean(m2)))/length(m1),sum((m1-mean(m1))*(m4-mean(m4)))/length(m1),sum((m1-mean(m1))*(m8-mean(m8)))/length(m1),sum((m1-mean(m1))*(m15-mean(m15)))/length(m1),sum((m1-mean(m1))*(m30-mean(m30)))/length(m1),sum((m1-mean(m1))*(mt-mean(mt)))/length(m1),sum((m1-mean(m1))*(m2-mean(m2)))/length(m1),sum((m2-mean(m2))^2)/length(m2),sum((m2-mean(m2))*(m4-mean(m4)))/length(m2),sum((m2-mean(m2))*(m8-mean(m8)))/length(m2),sum((m2-mean(m2))*(m15-mean(m15)))/length(m2),sum((m2-mean(m2))*(m30-mean(m30)))/length(m2),sum((m2-mean(m2))*(mt-mean(mt)))/length(m2),sum((m1-mean(m1))*(m4-mean(m4)))/length(m1),sum((m2-mean(m2))*(m4-mean(m4)))/length(m2),sum((m4-mean(m4))^2)/length(m4),sum((m4-mean(m4))*(m8-mean(m8)))/length(m4),sum((m4-mean(m4))*(m15-mean(m15)))/length(m4),sum((m4-mean(m4))*(m30-mean(m30)))/length(m4),sum((m4-mean(m4))*(mt-mean(mt)))/length(m4),sum((m1-mean(m1))*(m8-mean(m8)))/length(m1),sum((m2-mean(m2))*(m8-mean(m8)))/length(m2),sum((m4-mean(m4))*(m8-mean(m8)))/length(m4),sum((m8-mean(m8))^2)/length(m8),sum((m8-mean(m8))*(m15-mean(m15)))/length(m8),sum((m8-mean(m8))*(m30-mean(m30)))/length(m8),sum((m8-mean(m8))*(mt-mean(mt)))/length(m8),sum((m1-mean(m1))*(m15-mean(m15)))/length(m1),sum((m2-mean(m2))*(m15-mean(m15)))/length(m2),sum((m4-mean(m4))*(m15-mean(m15)))/length(m4),sum((m8-mean(m8))*(m15-mean(m15)))/length(m8),sum((m15-mean(m15))^2)/length(m15),sum((m15-mean(m15))*(m30-mean(m30)))/length(m15),sum((m15-mean(m15))*(mt-mean(mt)))/length(m15),sum((m1-mean(m1))*(m30-mean(m30)))/length(m1),sum((m2-mean(m2))*(m30-mean(m30)))/length(m2),sum((m4-mean(m4))*(m30-mean(m30)))/length(m4),sum((m8-mean(m8))*(m30-mean(m30)))/length(m8),sum((m15-mean(m15))*(m30-mean(m30)))/length(m15),sum((m30-mean(m30))^2)/length(m30),sum((m30-mean(m30))*(mt-mean(mt)))/length(m30),sum((m1-mean(m1))*(mt-mean(mt)))/length(m1),sum((m2-mean(m2))*(mt-mean(mt)))/length(m2),sum((m4-mean(m4))*(mt-mean(mt)))/length(m4),sum((m8-mean(m8))*(mt-mean(mt)))/length(m8),sum((m15-mean(m15))*(mt-mean(mt)))/length(m15),sum((m30-mean(m30))*(mt-mean(mt)))/length(m30),sum((mt-mean(mt))^2)/length(mt)),nrow=7,ncol=7),7)
Sn
R<-matrix(c(cor(m1,m1),cor(m1,m2),cor(m1,m4),cor(m1,m8),cor(m1,m15),cor(m1,m30),cor(m1,mt),cor(m2,m1),cor(m2,m2),cor(m2,m4),cor(m2,m8),cor(m2,m15),cor(m2,m30),cor(m2,mt),cor(m4,m1),cor(m4,m2),cor(m4,m4),cor(m4,m8),cor(m4,m15),cor(m4,m30),cor(m4,mt),cor(m8,m1),cor(m8,m2),cor(m8,m4),cor(m8,m8),cor(m8,m15),cor(m8,m30),cor(m8,mt),cor(m15,m1),cor(m15,m2),cor(m15,m4),cor(m15,m8),cor(m15,m15),cor(m15,m30),cor(m15,mt),cor(m30,m1),cor(m30,m2),cor(m30,m4),cor(m30,m8),cor(m30,m15),cor(m30,m30),cor(m30,mt),cor(mt,m1),cor(mt,m2),cor(mt,m4),cor(mt,m8),cor(mt,m15),cor(mt,m30),cor(mt,mt)),nrow=7,ncol=7)
R
```

The correlations between all pairs of races are positive and at least moderately strong (*r* $\geq$ `r cor(m1,mt)`), but we can see the correlation is stronger between races with similar distances.

## Problem 1.18

```{r Problem 1.18, tidy=TRUE, tidy.opts=list(width.cutoff=68)}
ms1<-100/m1
ms2<-200/m2
ms4<-400/m4
ms8<-800/(m8*60)
ms15<-1500/(m15*60)
ms30<-3000/(m30*60)
mts<-42195/(mt*60)
x<-matrix(c(mean(ms1),mean(ms2),mean(ms4),mean(ms8),mean(ms15),mean(ms30),mean(mts)))
x
Sn<-round(matrix(c(sum((ms1-mean(ms1))^2)/length(ms1),sum((ms1-mean(ms1))*(ms2-mean(ms2)))/length(ms1),sum((ms1-mean(ms1))*(ms4-mean(ms4)))/length(ms1),sum((ms1-mean(ms1))*(ms8-mean(ms8)))/length(ms1),sum((ms1-mean(ms1))*(ms15-mean(ms15)))/length(ms1),sum((ms1-mean(ms1))*(ms30-mean(ms30)))/length(ms1),sum((ms1-mean(ms1))*(mts-mean(mts)))/length(ms1),sum((ms1-mean(ms1))*(ms2-mean(ms2)))/length(ms1),sum((ms2-mean(ms2))^2)/length(ms2),sum((ms2-mean(ms2))*(ms4-mean(ms4)))/length(ms2),sum((ms2-mean(ms2))*(ms8-mean(ms8)))/length(ms2),sum((ms2-mean(ms2))*(ms15-mean(ms15)))/length(ms2),sum((ms2-mean(ms2))*(ms30-mean(ms30)))/length(ms2),sum((ms2-mean(ms2))*(mts-mean(mts)))/length(ms2),sum((ms1-mean(ms1))*(ms4-mean(ms4)))/length(ms1),sum((ms2-mean(ms2))*(ms4-mean(ms4)))/length(ms2),sum((ms4-mean(ms4))^2)/length(ms4),sum((ms4-mean(ms4))*(ms8-mean(ms8)))/length(ms4),sum((ms4-mean(ms4))*(ms15-mean(ms15)))/length(ms4),sum((ms4-mean(ms4))*(ms30-mean(ms30)))/length(ms4),sum((ms4-mean(ms4))*(mts-mean(mts)))/length(ms4),sum((ms1-mean(ms1))*(ms8-mean(ms8)))/length(ms1),sum((ms2-mean(ms2))*(ms8-mean(ms8)))/length(ms2),sum((ms4-mean(ms4))*(ms8-mean(ms8)))/length(ms4),sum((ms8-mean(ms8))^2)/length(ms8),sum((ms8-mean(ms8))*(ms15-mean(ms15)))/length(ms8),sum((ms8-mean(ms8))*(ms30-mean(ms30)))/length(ms8),sum((ms8-mean(ms8))*(mts-mean(mts)))/length(ms8),sum((ms1-mean(ms1))*(ms15-mean(ms15)))/length(ms1),sum((ms2-mean(ms2))*(ms15-mean(ms15)))/length(ms2),sum((ms4-mean(ms4))*(ms15-mean(ms15)))/length(ms4),sum((ms8-mean(ms8))*(ms15-mean(ms15)))/length(ms8),sum((ms15-mean(ms15))^2)/length(ms15),sum((ms15-mean(ms15))*(ms30-mean(ms30)))/length(ms15),sum((ms15-mean(ms15))*(mts-mean(mts)))/length(ms15),sum((ms1-mean(ms1))*(ms30-mean(ms30)))/length(ms1),sum((ms2-mean(ms2))*(ms30-mean(ms30)))/length(ms2),sum((ms4-mean(ms4))*(ms30-mean(ms30)))/length(ms4),sum((ms8-mean(ms8))*(ms30-mean(ms30)))/length(ms8),sum((ms15-mean(ms15))*(ms30-mean(ms30)))/length(ms15),sum((ms30-mean(ms30))^2)/length(ms30),sum((ms30-mean(ms30))*(mts-mean(mts)))/length(ms30),sum((ms1-mean(ms1))*(mts-mean(mts)))/length(ms1),sum((ms2-mean(ms2))*(mts-mean(mts)))/length(ms2),sum((ms4-mean(ms4))*(mts-mean(mts)))/length(ms4),sum((ms8-mean(ms8))*(mts-mean(mts)))/length(ms8),sum((ms15-mean(ms15))*(mts-mean(mts)))/length(ms15),sum((ms30-mean(ms30))*(mts-mean(mts)))/length(ms30),sum((mts-mean(mts))^2)/length(mts)),nrow=7,ncol=7),7)
Sn
R<-matrix(c(cor(ms1,ms1),cor(ms1,ms2),cor(ms1,ms4),cor(ms1,ms8),cor(ms1,ms15),cor(ms1,ms30),cor(ms1,mts),cor(ms2,ms1),cor(ms2,ms2),cor(ms2,ms4),cor(ms2,ms8),cor(ms2,ms15),cor(ms2,ms30),cor(ms2,mts),cor(ms4,ms1),cor(ms4,ms2),cor(ms4,ms4),cor(ms4,ms8),cor(ms4,ms15),cor(ms4,ms30),cor(ms4,mts),cor(ms8,ms1),cor(ms8,ms2),cor(ms8,ms4),cor(ms8,ms8),cor(ms8,ms15),cor(ms8,ms30),cor(ms8,mts),cor(ms15,ms1),cor(ms15,ms2),cor(ms15,ms4),cor(ms15,ms8),cor(ms15,ms15),cor(ms15,ms30),cor(ms15,mts),cor(ms30,ms1),cor(ms30,ms2),cor(ms30,ms4),cor(ms30,ms8),cor(ms30,ms15),cor(ms30,ms30),cor(ms30,mts),cor(mts,ms1),cor(mts,ms2),cor(mts,ms4),cor(mts,ms8),cor(mts,ms15),cor(mts,ms30),cor(mts,mts)),nrow=7,ncol=7)
R
```

Just like in Problem 1.17, the correlations between all pairs of races are positive and at least moderately strong (*r* $\geq$ `r cor(ms1,mts)`) and the correlation is stronger between races with similar distances. We can see this matrix is very similar to the matrix in Problem 1.17.