---
title: "STAT 488: Multivariate Statistical Analysis — Homework 2"
author: "Charles Hwang"
date: "2/21/2022"
output: pdf_document
---
## Problem 2.1

See previous page for Problem 2.1(a).

```{r Problem 2.1(b)}
rm(list=ls())
x<-matrix(c(5,1,3))
y<-matrix(c(-1,3,1))
sqrt(x[1]^2+x[2]^2+x[3]^2)                                                   # Problem 2.1(b)(i)
acos((x[1]*y[1]+x[2]*y[2]+x[3]*y[3])/(sqrt(sum(x^2))*sqrt(sum(y^2))))*180/pi # Problem 2.1(b)(ii)
as.numeric(t(y)%*%x/(x[1]^2+x[2]^2+x[3]^2))*x                                # Problem 2.1(b)(iii)
```

See previous page for Problem 2.1(c).

## Problem 2.6

### Problem 2.6(a)

```{r Problem 2.6(a)}
matrix(c(9,-2,-2,6),nrow=2,ncol=2)==t(matrix(c(9,-2,-2,6),nrow=2,ncol=2))
```

Yes, the matrix **A** is symmetric.

### Problem 2.6(b)

**x**$'$**Ax**$=9x_1^2-4x_1x_2+6x_2^2>0$ (what we are trying to show)

$9x_1^2-4x_1x_2+\frac{4}{9}x_2^2+\frac{50}{9}x_2^2>0$

$(3x_1-\frac{2}{3}x_2)^2+\frac{50}{9}x_2^2>0$

$\frac{1}{9}(9x_1-2x_2)^2+\frac{50}{9}x_2^2>0$

$(9x_1-2x_2)^2+50x_2^2>0$

$(9x_1-2x_2)^2>0$ $|$ $50x_2^2>0.$ $\square$

## Problem 2.7

### Problem 2.7(a)

$|$**A**$-\lambda$**I**$|=0$

$(9-\lambda)(6-\lambda)-(-2)(-2)=0$

$\lambda^2-15\lambda+50=0$

$(\lambda-5)(\lambda-10)=0$

$\lambda=5,10$

**Ax**$=\lambda$**x**

$9x_1-2x_2=5x_1$

$-2x_1+6x_2=5x_2$

$2x_1=x_2$

We can see there are an infinite number of solutions. If we arbitrarily pick $x_1=1$ and $x_2=2$, we can see that **e**$'=[\frac{1}{\sqrt5},\frac{2}{\sqrt5}]=[$ `r t(matrix(c(1,2))/as.numeric(sqrt(t(matrix(c(1,2)))%*%matrix(c(1,2)))))`$]$.

$9x_1-2x_2=10x_1$

$-2x_1+6x_2=10x_2$

$x_1=-2x_2$

We can see there are an infinite number of solutions. If we arbitrarily pick $x_1=2$ and $x_2=-1$, we can see that **e**$'=[\frac{2}{\sqrt5},\frac{-1}{\sqrt5}]=[$ `r t(matrix(c(2,-1))/as.numeric(sqrt(t(matrix(c(2,-1)))%*%matrix(c(2,-1)))))`$]$.

**Answers**: $\lambda_1=5,\lambda_2=10,$ **e**$'_1=[\frac{1}{\sqrt5},\frac{2}{\sqrt5}],$ **e**$'_2=[\frac{2}{\sqrt5},\frac{-1}{\sqrt5}]$

### Problem 2.7(b)

**A**$=\lambda_1$**e**$_1$**e**$'_1+\lambda_2$**e**$_2$**e**$'_2=5$**e**$_1$**e**$'_1+10$**e**$_2$**e**$'_2$

```{r Problem 2.7}
e1<-5*matrix(c(1/sqrt(5),2/sqrt(5)))%*%t(matrix(c(1/sqrt(5),2/sqrt(5))))
e1
e2<-10*matrix(c(2/sqrt(5),-1/sqrt(5)))%*%t(matrix(c(2/sqrt(5),-1/sqrt(5))))
e2
# We can see that the sum of these matrices is A.
solve(matrix(c(9,-2,-2,6),nrow=2,ncol=2)) # Problem 2.7(c)
```

### Problem 2.7(d)

$|$**A**$^{-1}-\lambda$**I**$|=0$

$(0.12-\lambda)(0.18-\lambda)-(0.04)(0.04)=0$

$\lambda^2-0.3\lambda+0.02=0$

$(\lambda-0.1)(\lambda-0.2)=0$

$\lambda=0.1,0.2$

**A**$^{-1}$**x**$=\lambda$**x**

$0.12x_1+0.04x_2=0.1x_1$

$0.04x_1+0.18x_2=0.1x_2$

$0.02x_1=-0.04x_2$

We can see there are an infinite number of solutions. If we arbitrarily pick $x_1=2$ and $x_2=-1$, we can see that **e**$'=[\frac{2}{\sqrt5},\frac{-1}{\sqrt5}]=$ `r t(matrix(c(2,-1))/as.numeric(sqrt(t(matrix(c(2,-1)))%*%matrix(c(2,-1)))))`.

$0.12x_1+0.04x_2=0.2x_1$

$0.04x_1+0.18x_2=0.2x_2$

$0.04x_1=0.02x_2$

We can see there are an infinite number of solutions. If we arbitrarily pick $x_1=1$ and $x_2=2$, we can see that **e**$'=[\frac{2}{\sqrt5},\frac{-1}{\sqrt5}]=$ `r t(matrix(c(1,2))/as.numeric(sqrt(t(matrix(c(1,2)))%*%matrix(c(1,2)))))`.

**Answers**: $\lambda_1=0.1,\lambda_2=0.2,$ **e**$'_1=[\frac{2}{\sqrt5},\frac{-1}{\sqrt5}],$ **e**$'_2=[\frac{1}{\sqrt5},\frac{2}{\sqrt5}]$

Observe that the eigenvalues for **A**$^{-1}$ are the multiplicative inverse/reciprocal ($\lambda^{-1}=\frac{1}{\lambda}$) of the eigenvalues for **A** and that the eigenvectors are for **A**$^{-1}$ and **A** are the same.

## Problem 2.15

**x**$'$**Ax**$=3x_1^2-2x_1x_2+3x_2^2$

$3x_1^2-6x_1x_2+3x_2^2+4x_1x_2$

$3(x_1^2-2x_1x_2+x_2^2)+4x_1x_2$

$3(x_1-x_2)^2+4x_1x_2$

$3(x_1-x_2)^2>0$ $|$ $4x_1x_2>0.$

We can see that the quadratic form $3x_1^2-2x_1x_2+3x_2^2$ is definite positive.

## Problem 2.25

```{r Problem 2.25, tidy=TRUE, tidy.opts=list(width.cutoff=82)}
E<-matrix(c(25,-2,4,-2,4,1,4,1,9),nrow=3,ncol=3) # Problem 2.25(a)
matrix(c(E[1,1]/sqrt(E[1,1]*E[1,1]),E[1,2]/sqrt(E[1,1]*E[2,2]),E[1,3]/sqrt(E[1,1]*E[3,3]),E[2,1]/sqrt(E[2,2]*E[1,1]),E[2,2]/sqrt(E[2,2]*E[2,2]),E[2,3]/sqrt(E[2,2]*E[3,3]),E[3,1]/sqrt(E[3,3]*E[1,1]),E[3,2]/sqrt(E[3,3]*E[2,2]),E[3,3]/sqrt(E[3,3]*E[3,3])),nrow=3,ncol=3)
matrix(c(sqrt(E[1,1]),0,0,0,sqrt(E[2,2]),0,0,0,sqrt(E[3,3])),nrow=3,ncol=3)
matrix(c(sqrt(E[1,1]),0,0,0,sqrt(E[2,2]),0,0,0,sqrt(E[3,3])),nrow=3,ncol=3)%*%matrix(c(E[1,1]/sqrt(E[1,1]*E[1,1]),E[1,2]/sqrt(E[1,1]*E[2,2]),E[1,3]/sqrt(E[1,1]*E[3,3]),E[2,1]/sqrt(E[2,2]*E[1,1]),E[2,2]/sqrt(E[2,2]*E[2,2]),E[2,3]/sqrt(E[2,2]*E[3,3]),E[3,1]/sqrt(E[3,3]*E[1,1]),E[3,2]/sqrt(E[3,3]*E[2,2]),E[3,3]/sqrt(E[3,3]*E[3,3])),nrow=3,ncol=3)%*%matrix(c(sqrt(E[1,1]),0,0,0,sqrt(E[2,2]),0,0,0,sqrt(E[3,3])),nrow=3,ncol=3) # Problem 2.25(b)
```

We can see that this is the same as the original matrix.

## Problem 2.31

```{r Problem 2.31}
matrix(c(4,3))                                                                     # Problem 2.31(a)
as.numeric(matrix(c(1,-1),ncol=2)%*%matrix(c(4,3)))                                # Problem 2.31(b)
matrix(c(3,0,0,1),nrow=2,ncol=2)                                                   # Problem 2.31(c)
as.numeric(matrix(c(1,-1),ncol=2)%*%matrix(c(3,0,0,1),nrow=2,ncol=2)%*%t(matrix(c(1,-1),ncol=2)))#d
matrix(c(2,1))                                                                     # Problem 2.31(e)
matrix(c(2,0,-1,1),nrow=2,ncol=2)%*%matrix(c(2,1))                                 # Problem 2.31(f)
matrix(c(9,-2,-2,4),nrow=2,ncol=2)                                                 # Problem 2.31(g)
matrix(c(2,0,-1,1),nrow=2,ncol=2)%*%matrix(c(9,-2,-2,4),nrow=2,ncol=2)%*%t(matrix(c(2,0,-1,1),nrow=2,ncol=2)) # Problem 2.31(h)
matrix(c(2,1,2,0),nrow=2,ncol=2)                                                   # Problem 2.31(i)
matrix(c(1,-1),ncol=2)%*%matrix(c(2,1,2,0),nrow=2,ncol=2)%*%t(matrix(c(2,0,-1,1),nrow=2,ncol=2)) #j
```

## Problem 4.3

We can see that (b) $X_2$ and $X_3$, (c) $(X_1,X_2)$ and $X_3$, and (d) $\frac{X_1+X_2}{2}$ and $X_3$ are independent because their respective covariances Cov$(X_2,X_3)=\frac{1}{2}$Cov$(X_1,X_3)+\frac{1}{2}$Cov$(X_2,X_3)=$Cov$(X_1,X_3)=0$, which is a sufficient condition for independence from Result 4.5(a) on page 159 of the textbook.

We can see that (a) $X_1$ and $X_2$ and (e) $X_2$ and $X_2-\frac{5}{2}X_1-X_3$ are not independent because their respective covariances Cov$(X_1,X_2)=-2\neq0$ and Cov$(X_2,X_2-\frac{5}{2}X_1-X_3)=$ `r t(matrix(c(-2,5,0)))%*%(matrix(c(-2,5,0))-5/2*matrix(c(1,-2,0))-matrix(c(0,0,2)))` $\neq0$.

## Problem 4.4(a)

$\mu=E(3X_1-2X_2+X_3)=3E(X_1)-2E(X_2)+E(X_3)=3(2)-2(-3)+(1)=6+6+(1)=$

$\mu=13$

$\Sigma=$Var$(3X_1-2X_2+X_3)=$

$\Sigma=a^2$Var$(X_1)+b^2$Var$(X_2)+c^2$Var$(X_3)+2ab$Cov$(X_1,X_2)+2ac$Cov$(X_1,X_3)+2bc$Cov$(X_2,X_3)=$

$\Sigma=3^2$Var$(X_1)+2^2$Var$(X_2)+$Var$(X_3)+2(3)(-2)$Cov$(X_1,X_2)+2(3)(1)$Cov$(X_1,X_3)+2(-2)(1)$Cov$(X_2,X_3)=$

$\Sigma=9(1)+4(3)+(2)-12(1)+6(1)-4(2)=9+12+(2)-12+6-8=$

$\Sigma=9$

We can see the distribution of $3X_1-2X_2+X_3$ is $N_3(13,9)$.

## Problem 4.19

(a) $\chi_6^2$ (from Result 4.7(a) on page 163 of the textbook)

(b) $N_{20}(\mu,\frac{1}{20}\Sigma)$ and $N_{20}(0,\Sigma)$ (from (4-23) 1. on page 174 of the textbook)

(c) $W_p(V,19)$, where $W$ is the Wishart distribution (from (4-23) 2. on page 174 of the textbook)

## Problem 4.29

```{r Problem 4.29, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=71)}
N<-read.table("/Users/newuser/Desktop/Notes/Graduate/STAT 488 - Multivariate Statistical Analysis/T1-5.dat")$V5
O<-read.table("/Users/newuser/Desktop/Notes/Graduate/STAT 488 - Multivariate Statistical Analysis/T1-5.dat")$V6
m<-matrix(c(mean(N),mean(O))) # Problem 4.29(a)
S<-solve(matrix(c(cov(N,N),cov(N,O),cov(O,N),cov(O,O)),nrow=2,ncol=2))
d<-matrix(c(t(matrix(c(N[1],O[1]))-m)%*%S%*%(matrix(c(N[1],O[1]))-m),t(matrix(c(N[2],O[2]))-m)%*%S%*%(matrix(c(N[2],O[2]))-m),t(matrix(c(N[3],O[3]))-m)%*%S%*%(matrix(c(N[3],O[3]))-m),t(matrix(c(N[4],O[4]))-m)%*%S%*%(matrix(c(N[4],O[4]))-m),t(matrix(c(N[5],O[5]))-m)%*%S%*%(matrix(c(N[5],O[5]))-m),t(matrix(c(N[6],O[6]))-m)%*%S%*%(matrix(c(N[6],O[6]))-m),t(matrix(c(N[7],O[7]))-m)%*%S%*%(matrix(c(N[7],O[7]))-m),t(matrix(c(N[8],O[8]))-m)%*%S%*%(matrix(c(N[8],O[8]))-m),t(matrix(c(N[9],O[9]))-m)%*%S%*%(matrix(c(N[9],O[9]))-m),t(matrix(c(N[10],O[10]))-m)%*%S%*%(matrix(c(N[10],O[10]))-m),t(matrix(c(N[11],O[11]))-m)%*%S%*%(matrix(c(N[11],O[11]))-m),t(matrix(c(N[12],O[12]))-m)%*%S%*%(matrix(c(N[12],O[12]))-m),t(matrix(c(N[13],O[13]))-m)%*%S%*%(matrix(c(N[13],O[13]))-m),t(matrix(c(N[14],O[14]))-m)%*%S%*%(matrix(c(N[14],O[14]))-m),t(matrix(c(N[15],O[15]))-m)%*%S%*%(matrix(c(N[15],O[15]))-m),t(matrix(c(N[16],O[16]))-m)%*%S%*%(matrix(c(N[16],O[16]))-m),t(matrix(c(N[17],O[17]))-m)%*%S%*%(matrix(c(N[17],O[17]))-m),t(matrix(c(N[18],O[18]))-m)%*%S%*%(matrix(c(N[18],O[18]))-m),t(matrix(c(N[19],O[19]))-m)%*%S%*%(matrix(c(N[19],O[19]))-m),t(matrix(c(N[20],O[20]))-m)%*%S%*%(matrix(c(N[20],O[20]))-m),t(matrix(c(N[21],O[21]))-m)%*%S%*%(matrix(c(N[21],O[21]))-m),t(matrix(c(N[22],O[22]))-m)%*%S%*%(matrix(c(N[22],O[22]))-m),t(matrix(c(N[23],O[23]))-m)%*%S%*%(matrix(c(N[23],O[23]))-m),t(matrix(c(N[24],O[24]))-m)%*%S%*%(matrix(c(N[24],O[24]))-m),t(matrix(c(N[25],O[25]))-m)%*%S%*%(matrix(c(N[25],O[25]))-m),t(matrix(c(N[26],O[26]))-m)%*%S%*%(matrix(c(N[26],O[26]))-m),t(matrix(c(N[27],O[27]))-m)%*%S%*%(matrix(c(N[27],O[27]))-m),t(matrix(c(N[28],O[28]))-m)%*%S%*%(matrix(c(N[28],O[28]))-m),t(matrix(c(N[29],O[29]))-m)%*%S%*%(matrix(c(N[29],O[29]))-m),t(matrix(c(N[30],O[30]))-m)%*%S%*%(matrix(c(N[30],O[30]))-m),t(matrix(c(N[31],O[31]))-m)%*%S%*%(matrix(c(N[31],O[31]))-m),t(matrix(c(N[32],O[32]))-m)%*%S%*%(matrix(c(N[32],O[32]))-m),t(matrix(c(N[33],O[33]))-m)%*%S%*%(matrix(c(N[33],O[33]))-m),t(matrix(c(N[34],O[34]))-m)%*%S%*%(matrix(c(N[34],O[34]))-m),t(matrix(c(N[35],O[35]))-m)%*%S%*%(matrix(c(N[35],O[35]))-m),t(matrix(c(N[36],O[36]))-m)%*%S%*%(matrix(c(N[36],O[36]))-m),t(matrix(c(N[37],O[37]))-m)%*%S%*%(matrix(c(N[37],O[37]))-m),t(matrix(c(N[38],O[38]))-m)%*%S%*%(matrix(c(N[38],O[38]))-m),t(matrix(c(N[39],O[39]))-m)%*%S%*%(matrix(c(N[39],O[39]))-m),t(matrix(c(N[40],O[40]))-m)%*%S%*%(matrix(c(N[40],O[40]))-m),t(matrix(c(N[41],O[41]))-m)%*%S%*%(matrix(c(N[41],O[41]))-m),t(matrix(c(N[42],O[42]))-m)%*%S%*%(matrix(c(N[42],O[42]))-m)),nrow=6,ncol=7)
d
mean(d<=qchisq(0.5,2))        # Problem 4.29(b)
plot(sort(d),qchisq((1:42-0.5)/42,2),main="Problem 4.29(c) - Chi-Square Plot of Ordered Distances")
abline(0,1)
```

## Problem 4.39

```{r Problem 4.39(a-b)}
T<-read.table("/Users/newuser/Desktop/Notes/Graduate/STAT 488 - Multivariate Statistical Analysis/T4-6.dat",col.names=c("I","S","B","C","L","G","E"))[1:5]
library(MVN)
par(mfrow=c(2,3))                                                          # Problem 4.39(a)
qqnorm(T$I)
qqline(T$I)
qqnorm(T$S)
qqline(T$S)
qqnorm(T$B)
qqline(T$B)
qqnorm(T$C)
qqline(T$C)
qqnorm(T$L)
qqline(T$L)
r<-data.frame(cor(data.frame(qqnorm(T$I,plot.it=F)))[2],cor(data.frame(qqnorm(T$S,plot.it=F)))[2],cor(data.frame(qqnorm(T$B,plot.it=F)))[2],cor(data.frame(qqnorm(T$C,plot.it=F)))[2],cor(data.frame(qqnorm(T$L,plot.it=F)))[2])
names(r)<-names(T)<-c("Independence","Support","Benevolence","Conformity","Leadership")
row.names(r)<-"r_Q"
r
# An online table (https://www.itl.nist.gov/div898/handbook/eda/section3/eda3676.htm)
# of critical values of the normal probability plot correlation coefficient shows the
# critical value of r_Q for n = 130 is approximately r = 0.9897. As such, we reject H0
# for the independence, support, and leadership variables at the alpha = 0.05 level.
# There is sufficient evidence these three variables are not normally distributed. We
# fail to reject H0 for the benevolence and conformity variables at the alpha = 0.05
# level. There is insufficient evidence these two variables are not normally distributed.
mvn(T,univariatePlot="scatter")$multivariateNormality                      # Problem 4.39(b)
# We fail to reject H0 at the alpha = 0.05 level. There is insufficient
# evidence (p = 0.104286) that the multivariate normality assumption has been violated.
```
```{r Problem 4.39(b-c)}
mvn(T,multivariateOutlierMethod="quan",showOutliers=TRUE)$multivariateOutliers
# We can see from the quantile method that there are 12 observed statistical outliers in
# the data, with observation 60 having the greatest Mahalanobis Distance of d = 28.732.
mvn(sqrt(T[c("Independence","Support","Leadership")]))$univariateNormality # Problem 4.39(c)
# Since these three variables are count data, (4-33) 1. on page 192 of the textbook says
# the square root scale is the most appropriate transformation in this situation. We can see
# the p-values of the Anderson-Darling test for the independence and leadership variables
# are greater than alpha = 0.05, but the p-value of the Anderson-Darling test for the
# variable for support is not, warranting additional analysis for this support variable.
mvn(T[c("Independence","Support","Leadership")],bc=TRUE)$univariateNormality
mvn(T[c("Independence","Support","Leadership")],bc=TRUE)$BoxCoxPowerTransformation
# I first performed a Box-Cox transformation using the function's built-in argument,
# but the p-value of the Anderson-Darling test for the support variable was still
# less than alpha = 0.05. We can see from the output that for whatever reason,
# lambda = 1.0 for the support variable, which meant that no scaling or transformation
# was actually performed. We can also note that lambda = 0.5 for the independence and
# leadership variables, which is the same as the initial square root transformation.
```
```{r Problem 4.39(c)}
mvn(data.frame(T$I,(T$S^1.8-1)/1.8,T$L),univariatePlot="histogram",bc=TRUE)$univariateNormality
mvn(data.frame(T$I,(T$S^1.5-1)/1.5,T$L),univariateTest="SF",bc=TRUE)$univariateNormality
mvn(data.frame(T$I,(T$S^1.5-1)/1.5,T$L),univariateTest="SW",bc=TRUE)$univariateNormality
# I then proceeded to perform a Box-Cox transformation manually using (4-34) on page 193.
# After testing various values of lambda, I found that lambda = 1.8 maximized the p-value
# of the Anderson-Darling test for the support variable, but it was still less than
# alpha = 0.05. We can see the histogram for the support variable appears to be bimodal
# with a mode around 50, which the function may consider to be not normally distributed.
# I eventually tried using different arguments for the univariate normality tests (since
# the default argument is the Anderson-Darling test) and additional values of lambda. This
# resulted in p-values greater than alpha = 0.05 when using the Shapiro-Francia test and
# the Shapiro-Wilk test with which I am most familiar, which both had maximized p-values
# when using lambda = 1.5. These results illustrate how different the various univariate
# normality tests can vary in methodology, test statistic, p-value, and conclusion.
rS<-data.frame(0.9897,cor(data.frame(qqnorm(T$S,plot.it=F)))[2],cor(data.frame(qqnorm(sqrt(T$S),plot.it=F)))[2],cor(data.frame(qqnorm(log(T$S),plot.it=F)))[2],cor(data.frame(qqnorm((T$S^1.8-1)/1.8,plot.it=F)))[2],cor(data.frame(qqnorm((T$S^1.5-1)/1.5,plot.it=F)))[2])
names(rS)<-c("Critical","Original","Square Root","Natural Log","Box-Cox (l=1.8)","Box-Cox (l=1.5)")
row.names(rS)<-"r_Q"
rS
```

I calculated the $r_Q$ values for the various transformations, and after some experimentation, I noticed a Box-Cox transformation using $\lambda=1.5$, not $\lambda=1.8$, provided the greatest value for $r_Q$ for the support variable.

In conclusion, it appears a **square root** ($\sqrt{x_j}$) transformation is the most appropriate for the variables for independence and leadership and a **Box-Cox** transformation with $\lambda=1.8$ is the most appropriate for the variable for support when using the Shapiro-Wilk test for univariate normality.