---
title: "Homework6_Hwang"
author: "Charles Hwang"
date: "4/13/2022"
output: pdf_document
---
Charles Hwang

Dr. Perry

STAT 451-001

13 April 2022

## Problem 1

### Problem 1(a)

```{r Problem 1(a)}
rm(list=ls())
plot(70,40,xlim=c(41,89),ylim=c(0,100),pch=19,xlab="Temperature (°F)",ylab="Humidity",main="Problem 1(a) - Partition Plot of Predicted Ozone")
abline(v=66)                                              # Branch 1
abline(v=c(59,80),col="red")                              # Branch 2
segments(c(0,66),c(29,77),c(59,80),c(29,77),col="green")  # Branch 3
segments(c(48,66),c(29,60),c(48,80),c(110,60),col="blue") # Branch 4
text(49,13,"4.1")
text(43.5,65,"4.6")
text(53.5,65,"8.2")
text(62.5,50,"9.3")
text(73,30,"12")
text(73,69,"16")
text(73,90,"22")
text(85.5,50,"23")
```

### Problem 1(b)

We can see from the partition plot in Problem 1(a) that the predicted ozone concentration is **12** for a temperature of 70°F and humidity of 40.

## Problem 2

```{r Problem 2, warning=FALSE}
bl<-read.csv("/Users/newuser/Desktop/Notes/Graduate/STAT 451 - Nonparametric Statistical Methods/BankLoan.csv")
bl$Personal.Loan<-as.factor(bl$Personal.Loan)
library(rpart)
library(tree)
set.seed(1304,sample.kind="Rounding")                               # Problem 2(a)
s<-sample(1:nrow(bl),nrow(bl)*.7)
train<-bl[s,]
test<-bl[-s,]
rpart<-rpart(Personal.Loan~.,data=train,method="class")             # Problem 2(b)
summary(rpart)
plot(rpart,uniform=TRUE,main="Problem 2(b) - Classification Tree")
text(rpart,use.n=TRUE,all=TRUE,cex=0.7)
mean(predict(rpart,newdata=train,type="class")==train$Personal.Loan) # Problem 2(c)
# We can see the accuracy of this initial model is approximately 98.57143 percent
# and that there are eight terminal nodes included in the classification tree.
mean(predict(rpart,newdata=test,type="class")==test$Personal.Loan)   # Problem 2(d)
table(Predicted=predict(rpart,newdata=test,type="class"),Actual=test$Personal.Loan)
cv<-cv.tree(tree(Personal.Loan~.,data=train))                       # Problem 2(e)
plot(cv$size,cv$dev,type="o",xlab="Number of Terminal Nodes",ylab="Cross-Validation SSE",main="Problem 2(e)")
# I believe the optimal number of nodes is 5. After pruning the tree to 5, 6, and 8 nodes,
# I saw there was very little loss of accuracy when using 5 nodes compared to 6 or 8.
ptree<-prune.tree(tree(Personal.Loan~.,data=train),best=5)          # Problem 2(f)
mean(predict(ptree,newdata=train,type="class")==train$Personal.Loan)
mean(predict(ptree,newdata=test,type="class")==test$Personal.Loan)  # Problem 2(g)
table(Predicted=predict(ptree,newdata=test,type="class"),Actual=test$Personal.Loan)
```

We can see the accuracy of this pruned tree is approximately `r 100*mean(predict(ptree,newdata=test,type="class")==test$Personal.Loan)` percent, which indicates this model seems to be predicting the outcome variable ("Personal.Loan") quite well.

## Problem 3

```{r Problem 3, message=FALSE, warning=FALSE}
w<-read.csv("/Users/newuser/Desktop/Notes/Graduate/STAT 451 - Nonparametric Statistical Methods/wage.csv")
w$maritl<-as.factor(w$maritl)         # Problem 3(a)
w$race<-as.factor(w$race)
w$education<-as.factor(w$education)
w$region<-as.factor(w$region)
w$jobclass<-as.factor(w$jobclass)
w$health<-as.factor(w$health)
w$health_ins<-as.factor(w$health_ins)
set.seed(1304,sample.kind="Rounding") # Problem 3(b)
ws<-sample(1:nrow(w),nrow(w)*.7)
wtrain<-w[ws,]
wtest<-w[-ws,]
wtree<-rpart(logwage~.,data=wtrain)   # Problem 3(c)
plot(wtree,uniform=TRUE,main="Problem 3(c) - Regression Tree")
text(wtree,use.n=TRUE,all=TRUE,cex=0.8)
mean((predict(wtree,newdata=wtrain)-wtrain$logwage)^2)
mean((predict(wtree,newdata=wtest)-wtest$logwage)^2)
wp<-read.csv("/Users/newuser/Desktop/Notes/Graduate/STAT 451 - Nonparametric Statistical Methods/wagepredict.csv")
wp$maritl<-as.factor(wp$maritl)       # Problem 3(d)
wp$race<-as.factor(wp$race)
wp$education<-as.factor(wp$education)
wp$region<-as.factor(wp$region)
wp$jobclass<-as.factor(wp$jobclass)
wp$health<-as.factor(wp$health)
wp$health_ins<-as.factor(wp$health_ins)
library(car)
Export(as.data.frame(predict(wtree,newdata=wp)),"Charles Hwang WagePredictions.csv")
```