---
title: "STAT Project 2: Multivariate Analysis of COVID-19 Daily Cases, Deaths, and Hospitalizations in the City of Chicago"
author: "Charles Hwang"
date: "5/3/2022"
output: pdf_document
---
The dataset I chose for this project is the "COVID-19 Daily Cases, Deaths, and Hospitalizations" dataset from the Chicago Data Portal. I chose this dataset because it is updated daily from the City of Chicago's public dashboard and all variables are numeric making it good for quantitative analysis. The dataset includes data for the 785 calendar days from March 1, 2020 through April 24, 2022.[^1] There are 58 variables in the dataset: (1) date; (2) total cases, (3) deaths, and (4) hospitalizations; (5-13) cases for ages 0-17, 18-29, 30-39, 40-49, 50-59, 60-69, 70-79, 80+, and unknown; (14-16) cases for female, male, and unknown; (17-22) cases for Latinx, Asian, black, white, other, and unknown races[^2]; and (23-40) the same groups for deaths and (41-58) hospitalizations.

I read the data into `R` and changed the date variable from character to the `Date` format. I renamed the variables to remove unreadable characters (hyphens) and for convenience and checked the total number of "unknown" values for each category and group (**Figure 1**). Over 96 percent (82,916) of "unknown" values were for race which suggests there may have been trouble or confusion in entering this data. Since the data are time-series, I figured it would be helpful to plot daily (**Figure 2a**) and cumulative (**Figure 3a**) cases, hospitalizations, and deaths over time and the area under the curves for deaths, and since cases are far greater than both hospitalizations and deaths, I plotted the same without them as insets of the original plots (**Figures 2b and 3b**). I also plotted the three variables for each age, gender, and race group[^3], which was not too interesting since the graphs all looked the same visually even with different *y*-axes (**Figures 2c through 2e**), but when I plotted the cumulative values with *all groups* on a single graph I was able to more easily compare them. The "0-17" age group had the most cases, generally decreasing with older age groups, but the opposite was true for deaths (**Figure 3c**). The "female" gender group had more cases but "male" had more hospitalizations and deaths (**Figure 3d**). Lastly, "Latinx" had the most cases, but "black" had the most hospitalizations and deaths of any race group by far (**Figure 3e**).

I then computed the correlation and covariance matrices for cases, hospitalizations, and deaths (**Figures 4a and 4b**) and found there are moderately strong positive correlations between the pairings of cases and hospitalizations and hospitalizations and deaths and a weak correlation between cases and deaths. There is a lot of covariance between cases and hospitalizations and considerable covariance between the other two pairs. I also computed correlation and covariance matrices for cases, hospitalizations, and deaths between age, gender, and race for inter-matrix analysis. There is strong to very strong positive correlation for cases between all age groups except unknown and moderate to strong positive correlation for hospitalizations between the same (**Figures 4c and 4d**). There are varying levels of positive correlation for deaths (**Figure 4e**), with no correlation when comparing the "0-17" age group to other groups rising progressively to moderately strong positive correlation when doing the same for the "80+" age group. The covariance matrices showed similar trends (**Figures 4f through 4h**), with very high covariance when comparing cases for the "0-17" age group to other groups decreasing progressively to relatively moderate covariance when doing the same for the "80+" age group. Hospitalizations and deaths showed the reverse trend with less covariance involved and nearly no covariance among many pairs of age groups for death. All matrices had little to no covariance for "unknown" age.

Since there were only three levels, I merged cases, hospitalizations, and deaths between gender onto the same matrix and interpreted the three $3\times3$ "submatrices" along the diagonal (think of a *Sudoku* grid). I found there was strong to very strong positive correlation between gender groups among all three variables (**Figure 4j**) and nearly perfect positive correlation between female and male cases ($r=0.99672$). Similar to age, the covariance matrix (**Figure 4k**) showed very high, relatively moderate, and little covariance between gender groups for for the three variables. There were no deaths of either unknown age or gender. Lastly, the matrices between race were similar to age without the progressive change among groups (**Figures 4m through 4s**), except for correlation for deaths, which were moderately positively correlated except for the other and unknown groups which had little to no correlation. I did not see any particularly noteworthy results between any two race groups. Overall, it appeared the results of all the matrices made sense intuitively. I was a bit intrigued at how extreme (both close to 1 and 0) some of the correlation coefficients were, but I figured this may be due to the robust sample size.

I performed principal component analysis (PCA) on total cases, hospitalizations, and deaths and found that the cases variable accounted for over 99.9 percent of the variance (**Figure 5a**). Looking at the signs of the principal components, the first component is a weighted sum, the second component compared cases to hospitalizations and deaths, and the third component compared hospitalizations to cases and deaths. Since there were only three variables, I realized it was only possible for components to be either weighted sums or comparing one variable to the other two. Since the basis of this project is multivariate analysis, I felt it was also meaningful to perform PCA on these variables individually between age and race categories.[^4] The **first component for all six analyses** was simply a **weighted sum** of the categories, and all variance proportions progressively decreased in order of input (age increasing or race decreasing by population). For cases and deaths between age groups, there was no practical interpretation as variables were not split contiguously and there was no intuitive comparison. The "0-17" age group accounted for over 95.2 percent of the variance in cases (**Figure 5b**) and nearly 80 percent of the variance in deaths (**Figure 5d**). For hospitalizations between age groups (**Figure 5c**), the third component compared those 0-69 with those 70+ and the fourth was a "60-69" (and unknown) component, but interpretations of other components are complicated for the same reason. The "0-17" age group again accounted for over 86.5 percent of the variance.

For cases between race groups (**Figure 5e**), the second component was a "Latinx" component, the third compared "white", "Asian", and "other" to the inverse, the fourth component compared "white", "black", and "Latinx" (the three most populous race groups) to the inverse, the fifth was a "Latinx"/"other" component, and the sixth compared "black", "Asian", and "other" (the three least populous race groups in the dataset) to the inverse. "White" accounted for over 91.6 percent of the variance. For hospitalizations between race groups (**Figure 5f**), the second component was a "black" component, the third was a "black"/"Latinx" component, and the fourth through sixth components compared the $\binom32=3$ possible pairs of "Asian", "other", and "unknown" to their inverses. "White" again accounted for nearly 90 percent of the variance. Lastly, for deaths between race groups (**Figure 5g**), the second component was a "black" component, the third compared "black", "Latinx", and "Asian" (the three most populous non-white race groups) to the inverse, the fourth was a "Asian"/"unknown" component, the fifth compared "white", "black", and "Latinx" to the inverse, and the sixth compared "white", "Asian", and "other" to the inverse. "White" accounted for over 80.2 percent of the variance. Looking at the cumulative plots (**Figures 3c and 3e**), these components made sense intuitively and accurately reflect the importance and proportions of each group.

I could not perform multivariate analysis of variance (MANOVA) because the data are time-series which violates the randomness and independence assumptions, and additionally several variables exhibit multicollinearity and do not appear to be multivariate normal. Thus, I performed factor analysis on the same six subsets, again excluding gender because I did not feel it was practical to interpret the factors of combined variables. For each subset, I chose the largest possible *m* that satisfies $(p-m)^2-p-m\geq0$ (where *p* is the number of variables in the subset) then performed an additional analysis with a smaller value of *m* after reviewing the initial factors (and the *p*-value from the accompanying $\chi^2$ goodness-of-fit test, if applicable). Of course, the selection of *m* is subjective, but I did not feel that testing every value of *m* would be useful so I just chose the maximum and "best" values. I also verified that $\Sigma=$ **LL**$'+\Psi$ (Example 9.1, pages 484-485) and wrote the code to produce factor scores from the weighted least squares and regression methods for each analysis, but did not print the results because I felt it would be impractical to include in the Appendix and not very meaningful to the overall analysis.

For both analyses of cases and hospitalizations between age groups (**Figures 6a through 6d**), the first factor was for old age and the second was for young age (as the loadings were in increasing and decreasing orders respectively), and the analyses for cases additionally had third and fourth factors for those 0-17. For deaths between age groups, the first factor was for old age, but the others were difficult to interpret as they were not in the order of age groups and there was no intuitive comparison (**Figures 6e and 6f**). For both analyses of cases between race groups (**Figure 6g and 6h**), the first factor was for "white", "black", "other", and "unknown" and the second factor was for "Latinx", and the $m=3$ analysis additionally had an "Asian" factor. For hospitalizations between race groups ($m=3$), the first factor was for "white" and "black", the second was for "unknown", and the third was for "Latinx" (**Figure 6j**). The $m=2$ analysis had a general factor (every group except "unknown") and a "white", "black", and "unknown" factor (**Figure 6k**). Lastly, for both analyses of deaths between race groups (**Figure 6m and 6n**), the first factor was for "Latinx", "black", and "white", and the $m=3$ analysis additionally had "Asian" and "white" factors. These factors also make sense intuitively based on the previous results of the plots and PCA.

Additional multivariate analyses like clustering, canonical correlation analysis (CCA), etc. could be performed if I had more time with this project. I could also perform time-series analysis incorporating multivariate topics like multiple regression models (Section 7.10, page 413) if I had more experience in these subjects (I believe this is the first university project for which I have used time-series data). Although the gender groups appeared to be very highly correlated, there could be additional analysis using this type of model in place of PCA and factor analysis. It does not appear to be possible with this dataset, but I would also be interested in investigations of the interactions between age, gender, and race and compare the three outcome variables for various sets of groups. If truly motivated, I could treat the unknown values (**Figure 1**) as missing data and use multiple imputation by chained equations (MICE), machine learning, or some similar method to predict the groups they belonged to. Either way, I believe I was able to rectify the mistakes of dataset selection from the previous project by vetting and choosing a robust dataset which made it easier to manipulate and interpret. I had some initial doubts I would be able to successfully analyze the dataset because it was so recent and may be more complex, but the variables were subsetted by age, gender, and race which was quite helpful. There did not appear to be any problems with performing any of the analyses.

Overall, I felt this was a good application of multivariate statistical methods. I found it interesting that the earliest date in the dataset was March 1, 2020, as the earliest recorded deaths in the United States were in January 2020 and the earliest recorded cases were in 2019. After some brief research, it appears March 1, 2020 was when the City of Chicago began recording these data or is at least the extent to which they have made them publicly available. I was able to choose a relevant and interesting dataset and perform thorough multivariate analyses using the topics learned in the second half of the course.

[^1]: I removed data from April 25 through April 29, 2022, the date I accessed and downloaded the dataset on, because hospitalization data was not yet available. I also removed another row because the date was missing.

[^2]: For all analyses, I rearranged the columns in the order of cases, hospitalizations, and deaths as that is the order of increasing severity and decreasing quantity. Similarly, I rearranged the race groups in the order of "white", "black", "Latinx", "Asian", "other", and "unknown" in the analysis, as that is the order of decreasing population based on traditional population estimates in the city of Chicago. There should be no change in the results. Although "Latinx" have outpopulated "black" in recent years, I felt it was easier for visual comparison and interpretation to list "black" in the second column.

[^3]: I excluded the "unknown" groups from race and gender for lack of data and convenience of fitting them in 2 x 4 and 2 x 1 output windows. I tried plotting all groups on a single graph like for the cumulative data, but the groups are very similar and the lines overlapped for each category, making it nearly impossible to reasonably interpret. I also tried plotting base-2 logarithm-transformed cumulative data, but the transformation appeared to be too extreme and unsuitable for the data (since there are multiple "peaks" or "surges").

[^4]: I excluded gender because there are only three levels and the same issue as before would arise. I also did not feel it was practical to interpret components if the three variables were combined.

## Appendix

### Figure 1

```{r Figure 1}
rm(list=ls())
C<-read.csv("/Users/newuser/Desktop/Notes/Graduate/STAT 488 - Multivariate Statistical Analysis/COVID-19_Daily_Cases__Deaths__and_Hospitalizations.csv",header=TRUE)
C$Date<-as.Date(C$Date,"%m/%d/%Y")
C<-na.omit(C[order(C$Date),])
row.names(C)<-1:nrow(C)
names(C)<-c("Date","C","D","H","C17","18C29","30C39","40C49","50C59","60C69","70C79","80C","C?A","CF","CM","C?G","CL","CA","CB","CW","Co","C?R","D17","18D29","30D39","40D49","50D59","60D69","70D79","80D","D?A","DF","DM","D?G","DL","DA","DB","DW","Do","D?R","H17","18H29","30H39","40H49","50H59","60H69","70H79","80H","H?A","HF","HM","H?G","HL","HA","HB","HW","Ho","H?R")
sapply(C[,c("C?A","H?A","D?A","C?G","H?G","D?G","C?R","H?R","D?R")],sum) # "Unknown"
```

### Figure 2a

```{r Figure 2a}
plot(C$Date,C$C,type="l",xlab="",ylab="",las=2,xaxt="n",main="COVID-19 Cases, Hospitalizations, and Deaths in Chicago")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$H,type="l",col="blue")
points(C$Date,C$D,type="l",col="red")
polygon(C$Date,C$D,col="orange",lty=0)
legend("topleft",c("Confirmed Cases","Hospitalizations","Confirmed Deaths"),col=c("black","blue","red"),lty=1)
```

### Figure 2b

```{r Figure 2b}
plot(C$Date,C$H,type="l",xlab="",ylab="",las=2,col="blue",xaxt="n",main="COVID-19 Hospitalizations and Deaths in Chicago")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$D,type="l",col="red")
segments(as.numeric(C$Date)[1],0,as.numeric(C$Date)[nrow(C)])
polygon(C$Date,C$D,col="orange",lty=0)
legend("topleft",c("Hospitalizations","Confirmed Deaths"),col=c("blue","red"),lty=1)
```

### Figure 2c

```{r Figure 2c}
par(mfrow=c(2,4))
plot(C$Date,C$C17,type="l",xlab="",ylab="",las=2,xaxt="n",main="Ages 0-17")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$H17,type="l",col="blue")
points(C$Date,C$D17,type="l",col="red")
plot(C$Date,C$`18C29`,type="l",xlab="",ylab="",las=2,xaxt="n",main="Ages 18-29")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$`18H29`,type="l",col="blue")
points(C$Date,C$`18D29`,type="l",col="red")
plot(C$Date,C$`30C39`,type="l",xlab="",ylab="",las=2,xaxt="n",main="Ages 30-39")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$`30H39`,type="l",col="blue")
points(C$Date,C$`30D39`,type="l",col="red")
plot(C$Date,C$`40C49`,type="l",xlab="",ylab="",las=2,xaxt="n",main="Ages 40-49")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$`40H49`,type="l",col="blue")
points(C$Date,C$`40D49`,type="l",col="red")
plot(C$Date,C$`50C59`,type="l",xlab="",ylab="",las=2,xaxt="n",main="Ages 50-59")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$`50H59`,type="l",col="blue")
points(C$Date,C$`50D59`,type="l",col="red")
plot(C$Date,C$`60C69`,type="l",xlab="",ylab="",las=2,xaxt="n",main="Ages 60-69")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$`60H69`,type="l",col="blue")
points(C$Date,C$`60D69`,type="l",col="red")
plot(C$Date,C$`70C79`,type="l",xlab="",ylab="",las=2,xaxt="n",main="Ages 70-79")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$`70H79`,type="l",col="blue")
points(C$Date,C$`70D79`,type="l",col="red")
plot(C$Date,C$`80C`,type="l",xlab="",ylab="",las=2,xaxt="n",main="Ages 80+")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$`80H`,type="l",col="blue")
points(C$Date,C$`80D`,type="l",col="red")
```

### Figure 2d

```{r Figure 2d}
par(mfrow=c(2,1))
plot(C$Date,C$CF,type="l",xlab="",ylab="",las=2,xaxt="n",main="Female COVID-19 Cases, Hospitalizations, and Deaths in Chicago")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$HF,type="l",col="blue")
points(C$Date,C$DF,type="l",col="red")
plot(C$Date,C$CM,type="l",xlab="",ylab="",las=2,xaxt="n",main="Male COVID-19 Cases, Hospitalizations, and Deaths in Chicago")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$HM,type="l",col="blue")
points(C$Date,C$DM,type="l",col="red")
```

### Figure 2e

```{r Figure 2e}
par(mfrow=c(2,3))
plot(C$Date,C$CW,type="l",xlab="",ylab="",las=2,xaxt="n",main="White")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$HW,type="l",col="blue")
points(C$Date,C$DW,type="l",col="red")
legend("topleft",c("Cases","Hospitalizations","Deaths"),cex=0.85,col=c("black","blue","red"),lty=1)
plot(C$Date,C$CB,type="l",xlab="",ylab="",las=2,xaxt="n",main="Black")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$HB,type="l",col="blue")
points(C$Date,C$DB,type="l",col="red")
plot(C$Date,C$CL,type="l",xlab="",ylab="",las=2,xaxt="n",main="Latinx")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$HL,type="l",col="blue")
points(C$Date,C$DL,type="l",col="red")
plot(C$Date,C$CA,type="l",xlab="",ylab="",las=2,xaxt="n",main="Asian")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$HA,type="l",col="blue")
points(C$Date,C$DA,type="l",col="red")
plot(C$Date,C$Co,type="l",xlab="",ylab="",las=2,xaxt="n",main="Other")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$Ho,type="l",col="blue")
points(C$Date,C$Do,type="l",col="red")
plot(C$Date,C$`C?R`,type="l",xlab="",ylab="",las=2,xaxt="n",main="Unknown")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,C$`H?R`,type="l",col="blue")
points(C$Date,C$`D?R`,type="l",col="red")
```

### Figure 3a

```{r Figure 3a}
plot(C$Date,cumsum(C$C),type="l",xlab="",ylab="",xaxt="n",yaxt="n",main="Cumulative Chicago COVID-19 Cases, Hospitalizations, & Deaths")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
axis(2,axTicks(2),format(axTicks(2),scientific=FALSE),las=2)
points(C$Date,cumsum(C$H),type="l",col="blue")
points(C$Date,cumsum(C$D),type="l",col="red")
polygon(c(C$Date,max(C$Date)),c(cumsum(C$D),0),col="orange",lty=0)
legend("topleft",c("Cumulative Confirmed Cases","Cumulative Hospitalizations","Cumulative Confirmed Deaths"),col=c("black","blue","red"),lty=1)
```

### Figure 3b

```{r Figure 3b}
plot(C$Date,cumsum(C$H),type="l",xlab="",ylab="",las=2,col="blue",xaxt="n",main="Cumulative Chicago COVID-19 Hospitalizations and Deaths")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,cumsum(C$D),type="l",col="red")
polygon(c(C$Date,max(C$Date)),c(cumsum(C$D),0),col="orange",lty=0)
legend("topleft",c("Cumulative Hospitalizations","Cumulative Confirmed Deaths"),col=c("blue","red"),lty=1)
```

### Figure 3c

```{r Figure 3c}
par(mfrow=c(1,3))
plot(C$Date,cumsum(C$`18C29`),type="l",xlab="",ylab="",las=2,xaxt="n",main="Cumulative Cases")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,cumsum(C$C17),type="l",col="skyblue")
points(C$Date,cumsum(C$`30C39`),type="l",col="blue")
points(C$Date,cumsum(C$`40C49`),type="l",col="green4")
points(C$Date,cumsum(C$`50C59`),type="l",col="green")
points(C$Date,cumsum(C$`60C69`),type="l",col="yellow")
points(C$Date,cumsum(C$`70C79`),type="l",col="orange")
points(C$Date,cumsum(C$`80C`),type="l",col="red")
points(C$Date,cumsum(C$`C?A`),type="l",col="gray")
legend("topleft",c("Ages 0-17","Ages 18-29","Ages 30-39","Ages 40-49","Ages 50-59","Ages 60-69","Ages 70-79","Ages 80+","Unknown"),col=c("black","skyblue","blue","green4","green","yellow","orange","red","gray"),lty=1)
plot(C$Date,cumsum(C$`60H69`),col="yellow",type="l",xlab="",ylab="",las=2,xaxt="n",main="Cumulative Hospitalizations")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,cumsum(C$H17),type="l",col="skyblue")
points(C$Date,cumsum(C$`18H29`),type="l")
points(C$Date,cumsum(C$`30H39`),type="l",col="blue")
points(C$Date,cumsum(C$`40H49`),type="l",col="green4")
points(C$Date,cumsum(C$`50H59`),type="l",col="green")
points(C$Date,cumsum(C$`70H79`),type="l",col="orange")
points(C$Date,cumsum(C$`80H`),type="l",col="red")
points(C$Date,cumsum(C$`H?A`),type="l",col="gray")
legend("topleft",c("0-17","18-29","30-39","40-49","50-59","60-69","70-79","80+","Unk."),col=c("black","skyblue","blue","green4","green","yellow","orange","red","gray"),lty=1)
plot(C$Date,cumsum(C$`80D`),col="red",type="l",xlab="",ylab="",las=2,xaxt="n",main="Cumulative Deaths")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,cumsum(C$D17),type="l",col="skyblue")
points(C$Date,cumsum(C$`18D29`),type="l")
points(C$Date,cumsum(C$`30D39`),type="l",col="blue")
points(C$Date,cumsum(C$`40D49`),type="l",col="green4")
points(C$Date,cumsum(C$`50D59`),type="l",col="green")
points(C$Date,cumsum(C$`60D69`),type="l",col="yellow")
points(C$Date,cumsum(C$`70D79`),type="l",col="orange")
points(C$Date,cumsum(C$`D?A`),type="l",col="gray")
legend("topleft",c("0-17","18-29","30-39","40-49","50-59","60-69","70-79","80+","Unk."),col=c("black","skyblue","blue","green4","green","yellow","orange","red","gray"),lty=1)
```

### Figure 3d

```{r Figure 3d}
par(mfrow=c(1,2))
plot(C$Date,cumsum(C$CF),type="l",xlab="",ylab="",las=2,xaxt="n",main="Cumulative Cases, Gender")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,cumsum(C$CM),type="l",col="red")
points(C$Date,cumsum(C$`C?G`),type="l",col="gray")
legend("topleft",c("Female","Male","Unknown"),col=c("black","red","gray"),lty=1)
plot(C$Date,cumsum(C$HM),col="red",type="l",xlab="",ylab="",las=2,xaxt="n",main="Cum. Hosp./Deaths, Gender")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,cumsum(C$HF),type="l")
points(C$Date,cumsum(C$`H?G`),type="l",col="gray")
points(C$Date,cumsum(C$DM),type="l",col="orange")
points(C$Date,cumsum(C$DF),type="l",col="blue")
points(C$Date,cumsum(C$`D?G`),type="l",col="gray")
legend("topleft",c("H (M)","H (F)","D (M)","D (F)","Unk."),col=c("red","black","orange","blue","gray"),lty=1)
```

### Figure 3e

```{r Figure 3e}
par(mfrow=c(1,3))
plot(C$Date,cumsum(C$CL),type="l",xlab="",ylab="",las=2,xaxt="n",main="Cumulative Cases")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,cumsum(C$CW),type="l",col="tan")
points(C$Date,cumsum(C$CB),type="l",col="brown")
points(C$Date,cumsum(C$CA),type="l",col="yellow")
points(C$Date,cumsum(C$Co),type="l",col="blue")
points(C$Date,cumsum(C$`C?R`),type="l",col="red")
legend("topleft",c("White","Black","Latinx","Asian","Other","Unknown"),col=c("tan","brown","black","yellow","blue","red"),lty=1)
plot(C$Date,cumsum(C$HB),col="brown",type="l",xlab="",ylab="",las=2,xaxt="n",main="Cumulative Hospitalizations")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,cumsum(C$HW),type="l",col="tan")
points(C$Date,cumsum(C$HL),type="l")
points(C$Date,cumsum(C$HA),type="l",col="yellow")
points(C$Date,cumsum(C$Ho),type="l",col="blue")
points(C$Date,cumsum(C$`H?R`),type="l",col="red")
legend("topleft",c("White","Black","Latinx","Asian","Other","Unknown"),col=c("tan","brown","black","yellow","blue","red"),lty=1)
plot(C$Date,cumsum(C$DB),col="brown",type="l",xlab="",ylab="",las=2,xaxt="n",main="Cumulative Deaths")
axis.Date(1,at=seq(min(C$Date),max(C$Date),by="1 mon"),format="%m-%y",las=2)
points(C$Date,cumsum(C$DW),type="l",col="tan")
points(C$Date,cumsum(C$DL),type="l")
points(C$Date,cumsum(C$DA),type="l",col="yellow")
points(C$Date,cumsum(C$Do),type="l",col="blue")
points(C$Date,cumsum(C$`D?R`),type="l",col="red")
legend("topleft",c("White","Black","Latinx","Asian","Other","Unknown"),col=c("tan","brown","black","yellow","blue","red"),lty=1)
```

### Figure 4a

```{r Figure 4a}
cor(C[,c("C","H","D")])
```

### Figure 4b

```{r Figure 4b}
cov(C[,c("C","H","D")])
```

### Figure 4c

```{r Figure 4c}
CA<-C[,c("C17","18C29","30C39","40C49","50C59","60C69","70C79","80C","C?A")]
round(cor(CA),5)
```

### Figure 4d

```{r Figure 4d}
HA<-C[,c("H17","18H29","30H39","40H49","50H59","60H69","70H79","80H","H?A")]
round(cor(HA),5)
```

### Figure 4e

```{r Figure 4e}
DA<-C[,c("D17","18D29","30D39","40D49","50D59","60D69","70D79","80D","D?A")]
round(cor(DA),5)
```

### Figure 4f

```{r Figure 4f}
round(cov(CA),1)
```

### Figure 4g

```{r Figure 4g}
round(cov(HA),4)
```

### Figure 4h

```{r Figure 4h}
round(cov(DA),5)
```

### Figure 4j

```{r Figure 4j}
G<-C[,c("CF","CM","C?G","HF","HM","H?G","DF","DM","D?G")]
round(cor(G),5)
```

### Figure 4k

```{r Figure 4k}
round(cov(G),2)
```

### Figure 4m

```{r Figure 4m}
CR<-C[,c("CW","CB","CL","CA","Co","C?R")]
cor(CR)
```

### Figure 4n

```{r Figure 4n}
HR<-C[,c("HW","HB","HL","HA","Ho","H?R")]
cor(HR)
```

### Figure 4p

```{r Figure 4p}
DR<-C[,c("DW","DB","DL","DA","Do","D?R")]
cor(DR)
```

### Figure 4q

```{r Figure 4q}
cov(CR)
```

### Figure 4r

```{r Figure 4r}
cov(HR)
```

### Figure 4s

```{r Figure 4s}
cov(DR)
```

### Figure 5a

```{r Figure 5a}
prcomp(C[,c("C","H","D")])
prcomp(C[,c("C","H","D")])$sdev[1]^2/sum(prcomp(C[,c("C","H","D")])$sdev^2)
```

### Figure 5b

```{r Figure 5b}
prcomp(CA)
prcomp(CA)$sdev[1]^2/sum(prcomp(CA)$sdev^2)
```

### Figure 5c

```{r Figure 5c}
prcomp(HA)
prcomp(HA)$sdev[1]^2/sum(prcomp(HA)$sdev^2)
```

### Figure 5d

```{r Figure 5d}
DAa<-DA[,-9]
prcomp(DAa)
prcomp(DAa)$sdev[1]^2/sum(prcomp(DAa)$sdev^2)
```

### Figure 5e

```{r Figure 5e}
prcomp(CR)
prcomp(CR)$sdev[1]^2/sum(prcomp(CR)$sdev^2)
```

### Figure 5f

```{r Figure 5f}
prcomp(HR)
prcomp(HR)$sdev[1]^2/sum(prcomp(HR)$sdev^2)
```

### Figure 5g

```{r Figure 5g}
prcomp(DR)
prcomp(DR)$sdev[1]^2/sum(prcomp(DR)$sdev^2)
```

### Figure 6a

```{r Figure 6a}
factanal(CA,5)
```
```{r Figure 6aN, eval=FALSE}
round(factanal(CA,5)$loadings%*%t(factanal(CA,5)$loadings)+factanal(CA,5)$uniqueness,4)
cor(CA)
factanal(CA,5,scores="Bartlett")$scores
factanal(CA,5,scores="regression")$scores
```

### Figure 6b

```{r Figure 6b}
factanal(CA,4)
```
```{r Figure 6bN, eval=FALSE}
round(factanal(CA,4)$loadings%*%t(factanal(CA,4)$loadings)+factanal(CA,4)$uniqueness,4)
cor(CA)
factanal(CA,4,scores="Bartlett")$scores
factanal(CA,4,scores="regression")$scores
```

### Figure 6c

```{r Figure 6c}
factanal(HA,5)
```
```{r Figure 6cN, eval=FALSE}
round(factanal(HA,5)$loadings%*%t(factanal(HA,5)$loadings)+factanal(HA,5)$uniqueness,4)
cor(HA)
factanal(HA,4,scores="Bartlett")$scores
factanal(HA,4,scores="regression")$scores
```

### Figure 6d

```{r Figure 6d}
factanal(HA,2)
```
```{r Figure 6dN, eval=FALSE}
round(factanal(HA,2)$loadings%*%t(factanal(HA,2)$loadings)+factanal(HA,2)$uniqueness,4)
cor(HA)
factanal(HA,2,scores="Bartlett")$scores
factanal(HA,2,scores="regression")$scores
```

### Figure 6e

```{r Figure 6e}
factanal(DAa,4)
```
```{r Figure 6eN, eval=FALSE}
round(factanal(DAa,4)$loadings%*%t(factanal(DAa,4)$loadings)+factanal(DAa,4)$uniqueness,5)
cor(DAa)
factanal(DAa,4,scores="Bartlett")$scores
factanal(DAa,4,scores="regression")$scores
```

### Figure 6f

```{r Figure 6f}
factanal(DAa,1)
```
```{r Figure 6fN, eval=FALSE}
round(factanal(DAa,1)$loadings%*%t(factanal(DAa,1)$loadings)+factanal(DAa,1)$uniqueness,5)
cor(DAa)
factanal(DAa,1,scores="Bartlett")$scores
factanal(DAa,1,scores="regression")$scores
```

### Figure 6g

```{r Figure 6g}
factanal(CR,3)
```
```{r Figure 6gN, eval=FALSE}
factanal(CR,3)$loadings%*%t(factanal(CR,3)$loadings)+factanal(CR,3)$uniqueness
cor(CR)
factanal(CR,3,scores="Bartlett")$scores
factanal(CR,3,scores="regression")$scores
```

### Figure 6h

```{r Figure 6h}
factanal(CR,2)
```
```{r Figure 6hN, eval=FALSE}
factanal(CR,2)$loadings%*%t(factanal(CR,2)$loadings)+factanal(CR,2)$uniqueness
cor(CR)
factanal(CR,2,scores="Bartlett")$scores
factanal(CR,2,scores="regression")$scores
```

### Figure 6j

```{r Figure 6j}
factanal(HR,3)
```
```{r Figure 6jN, eval=FALSE}
factanal(HR,3)$loadings%*%t(factanal(HR,3)$loadings)+factanal(HR,3)$uniqueness
cor(HR)
factanal(HR,3,scores="Bartlett")$scores
factanal(HR,3,scores="regression")$scores
```

### Figure 6k

```{r Figure 6k}
factanal(HR,2)
```
```{r Figure 6kN, eval=FALSE}
factanal(HR,2)$loadings%*%t(factanal(HR,2)$loadings)+factanal(HR,2)$uniqueness
cor(HR)
factanal(HR,2,scores="Bartlett")$scores
factanal(HR,2,scores="regression")$scores
```

### Figure 6m

```{r Figure 6m}
factanal(DR,3)
```
```{r Figure 6mN, eval=FALSE}
factanal(DR,3)$loadings%*%t(factanal(DR,3)$loadings)+factanal(DR,3)$uniqueness
cor(DR)
factanal(DR,3,scores="Bartlett")$scores
factanal(DR,3,scores="regression")$scores
```

### Figure 6n

```{r Figure 6n}
factanal(DR,1)
```
```{r Figure 6nN, eval=FALSE}
factanal(DR,1)$loadings%*%t(factanal(DR,1)$loadings)+factanal(DR,1)$uniqueness
cor(DR)
factanal(DR,1,scores="Bartlett")$scores
factanal(DR,1,scores="regression")$scores
```

## Works Cited

| "COVID-19 Daily Cases, Deaths, and Hospitalizations." *Chicago Data Portal*, 24 April 2020,
|         <https://data.cityofchicago.org/d/naz8-j4nc>. Accessed 29 April 2022.