---
title: "Homework 3"
author: "Charles Hwang"
date: "6/23/2022"
output: pdf_document
---
Charles Hwang

Dr. Matthews

STAT 488-001

23 June 2022

## Problem 5.13

### Problem 5.13(b)

```{r Problem 5.13(b)}
rm(list=ls())
y<-c(16,9,10,13,19,20,18,17,35,55)
o<-c(58,90,48,57,103,57,86,112,273,64)
n<-y+o
mean(y/n)    # mean = alpha / (alpha+beta)
var(y/n) # variance = alpha*beta / ((alpha+beta)^2*(alpha+beta+1))
a<-2.6 # Solving for alpha and beta in Desmos and rounding to nearest tenth
b<-10.6
mp<-function(X,Y){a<-exp(Y)*exp(X)/(exp(X)+1)
  b<-exp(Y)/(exp(X)+1)
  10^230*(a+b)^(-2)*prod(beta(a+y,b+n-y)/beta(a,b))*exp(2*Y+X)/(exp(X)+1)^2} # Jacobian
X<-seq(-1.8,-0.9,length=50) # Setting bounds of contour plot via trial-and-error
Y<-seq(1.9,4.1,length=50)
z<-matrix(NA,length(Y),length(X))
for (i in 1:length(Y)){for (j in 1:length(X)){z[j,i]<-mp(X[j],Y[i])}}
par(mfrow=c(1,2))
contour(X,Y,z,xlab="ln(alpha / beta)",ylab="ln(alpha + beta)",main="Problem 5.13(b) - Contour Plot")
points(log(a/b),log(a+b),pch=19)
Xm<-Ym<-Xn<-Yn<-rep(NA,500)
set.seed(226)
for (k in 1:length(Xm)){Xm[k]<-sample(X,1,TRUE,apply(z,1,sum)/sum(apply(z,1,sum)))
  Ym[k]<-sample(Y,1,TRUE,z[which(Xm[k]==X),]/sum(z[which(Xm[k]==X),]))}
contour(X,Y,z,xlab="ln(alpha / beta)",ylab="ln(alpha + beta)",main="Draws from Marginal Posterior")
Xm<-Xm+runif(length(Xm),-diff(X)/2,diff(X)/2) # Adding random jitter
Ym<-Ym+runif(length(Ym),-diff(Y)/2,diff(Y)/2)
points(Xm,Ym)
points(mean(Xm),mean(Ym),pch=19,col="red") # Mean of samples
alpha<-exp(Ym)*(exp(Xm)/(exp(Xm)+1)) # Now drawing from joint posterior distribution
beta<-exp(Ym)/(exp(Xm)+1)
par(mfrow=c(1,1))
plot(alpha,beta,main="Problem 5.13(b) - 500 Draws from Joint Posterior Distribution")
abline(0,b/a) # Expected slope
par(mfrow=c(1,2))
hist(alpha)
hist(beta)
```

### Problem 5.13(c)

```{r Problem 5.13(c)}
library(ggplot2)
t<-matrix(NA,5000,length(y))
pd<-data.frame()
set.seed(226)
for (i in 1:length(y)){t[,i]<-rbeta(5000,alpha+y[i],beta+n[i]-y[i])
pd<-rbind(pd,data.frame(t[,i],rep(i,5000)))}
names(pd)<-c("theta","i")
par(mfrow=c(2,5))
hist(subset(pd,i==1)$theta,xlab="",ylab="",main="")
abline(v=(y/n)[1],col="red")
hist(subset(pd,i==2)$theta,xlab="",ylab="",main="")
abline(v=(y/n)[2],col="red")
hist(subset(pd,i==3)$theta,xlab="",ylab="",main="")
abline(v=(y/n)[3],col="red")
hist(subset(pd,i==4)$theta,xlab="",ylab="",main="")
abline(v=(y/n)[4],col="red")
hist(subset(pd,i==5)$theta,xlab="",ylab="",main="")
abline(v=(y/n)[5],col="red")
hist(subset(pd,i==6)$theta,xlab="",ylab="",main="")
abline(v=(y/n)[6],col="red")
hist(subset(pd,i==7)$theta,xlab="",ylab="",main="")
abline(v=(y/n)[7],col="red")
hist(subset(pd,i==8)$theta,xlab="",ylab="",main="")
abline(v=(y/n)[8],col="red")
hist(subset(pd,i==9)$theta,xlab="",ylab="",main="")
abline(v=(y/n)[9],col="red")
hist(subset(pd,i==10)$theta,xlab="",ylab="",main="")
abline(v=(y/n)[10],col="red")
par(mfrow=c(1,1)) # Plotting posterior means vs. raw proportions
tj<-sapply(split(pd$theta,pd$i),mean)
plot(y/n,tj,ylab="mean(theta_j)",main="Problem 5.13(c) - Posterior Means vs. Raw Proportions (y/n)")
abline(0,1)
```

The inferences from the posterior distributions for the $\theta_j$'s are generally centered around the raw proportion ($\frac{y_j}{n_j}$). This is expected as $\theta_j|\alpha,\beta,y\sim Beta(\alpha+y_j,\beta+n_j-y_j)$ and it is apparent that 5,000 samples is sufficient. The center of the posterior distribution for $\theta_{10}$ appears to be considerably less (`r mean(subset(pd,i==10)$theta)`) than the raw proportion $\frac{y_{10}}{n_{10}}=\frac{55}{55+64}=\frac{55}{119}\approx0.4621849$, but we can see from the data that this particular location has a far greater proportion of bicycles than the other locations (its raw proportion is more than twice the next highest, $\theta_1=\frac{y_1}{n_1}=\frac{16}{16+58}=\frac{16}{74}=\frac{8}{37}=0.\overline{216}$) and could reasonably be considered an outlier location.

### Problem 5.13(d)

```{r Problem 5.13(d)}
quantile(alpha/(alpha+beta),c(0.05/2,1-0.05/2))
```

### Problem 5.13(e)

```{r Problem 5.13(e)}
set.seed(226) # Sampling 100 draws from beta (variance) and binomial distributions
quantile(rbinom(100,100,rbeta(100,alpha,beta)),c(0.05/2,1-0.05/2))
```

I do not *fully* trust this posterior interval in application because this new location may be correlated with other locations or not fully independent, or conversely could be much different than the original locations. We also do not know from the problem whether the number of bicycles at this location follows a binomial distribution, but that is our best guess. The interval is rather conservative (wide), but additional information on the location may be needed to use it in practice.

## Problem 5.14

### Problem 5.14(b)

```{r Problem 5.14(b)}
mean(y/n)             # mean = alpha / beta
var(y/n)          # variance = alpha / beta^2
b<-mean(y/n)/var(y/n) # beta = mean / variance
a<-b*mean(y/n)       # alpha = beta * mean
mpd<-function(X,Y){a<-exp(Y)*exp(X)/(exp(X)+1)
  b<-exp(Y)/(exp(X)+1)
  10^-200*sqrt(a)*(b+1)*prod(b^a/(b+1)^(y+a)*gamma(y)/beta(a,y))*exp(2*Y+X)/(exp(X)+1)^2}
X<-seq(2.3,3.4,length=50) # Setting bounds of contour plot via trial-and-error
Y<-seq(1,3.6,length=50)
z<-matrix(NA,length(Y),length(X))
for (i in 1:length(Y)){for (j in 1:length(X)){z[j,i]<-mpd(X[j],Y[i])}}
par(mfrow=c(1,2))
contour(X,Y,z,xlab="ln(alpha / beta)",ylab="ln(alpha + beta)",main="Problem 5.14(b) - Contour Plot")
set.seed(226)
for (k in 1:length(Xn)){Xn[k]<-sample(X,1,TRUE,apply(z,1,sum)/sum(apply(z,1,sum)))
  Yn[k]<-sample(Y,1,TRUE,z[which(Xn[k]==X),]/sum(z[which(Xn[k]==X),]))}
contour(X,Y,z,xlab="ln(alpha / beta)",ylab="ln(alpha + beta)",main="Draws from Marginal Posterior")
Xn<-Xn+runif(length(Xn),-diff(X)/2,diff(X)/2) # Adding random jitter
Yn<-Yn+runif(length(Yn),-diff(Y)/2,diff(Y)/2)
points(Xn,Yn) # One extreme outlier at (~2.3,~1.5)
points(mean(Xn),mean(Yn),pch=19,col="red") # Mean of samples
```

### Problem 5.14(c)

Yes, the posterior density ($p(\theta,\alpha,\beta|y)\propto p(\alpha,\beta)\prod_{j=1}^J\frac{\beta^\alpha}{\Gamma(\alpha)}\theta_j^{y_j+\alpha-1}e^{-(\beta+1)\theta_j}$) appears to be integrable. We can see from the plots in problems 5.14(e) and 5.14(b) that the joint posterior and contours are finite and do not diverge. Any function can be integrated if you try hard enough and have enough computing power.

### Problem 5.14(e)

```{r Problem 5.14(e)}
alpha<-exp(Yn)*(exp(Xn)/(exp(Xn)+1))
beta<-exp(Yn)/(exp(Xn)+1)
plot(alpha,beta,main="Problem 5.14(e) - 500 Draws from Joint Posterior Distribution")
abline(0,0.05) # Unknown where the slope comes from?
```