---
title: "Homework 3"
author: "Charles Hwang"
date: "2/17/2022"
output: pdf_document
---
Charles Hwang

Dr. Whalen

STAT 410-001

17 February 2022

## Problem 3.1

The link function provides the proper function to relate the response variable. The identity link is the link function $g(\mu)=\mu$ and it is seldom used with a binomial distribution because the range of the identity link, $[-\infty,\infty]$, is different than the range of the binomial parameter, which is defined as being 0 or 1.

## Problem 3.2

### Problem 3.2a

"For every 10 years since 1904, the model states the probability of a given starting pitcher pitching a complete game decreases by approximately 6.62 percent."

### Problem 3.2b

$\hat{P}(Y=1)=0.6930-0.0662x=0.6930-0.0662(12)=-0.1014$

The prediction from the logistic regression model is clearly more plausible. The predition from the linear model is outside the range of possible probabilities $[0,1]$.

## Problem 3.5

### Problem 3.5a

```{r Problem 3.5a}
rm(list=ls())
hc<-read.table("http://users.stat.ufl.edu/~aa/cat/data/Crabs.dat",header=TRUE)
ci<-lm(y~weight,data=hc)
summary(ci)
```

Interpretation of *intercept*: "If the weight of a crab was 0 kilograms, the model states the probability of a crab having at least one satellite is approximately `r ci$coefficients["(Intercept)"]`" (which obviously does not make sense).

Interpretation of *coefficient for weight*: "For every 1 kilogram increase in weight of a crab, the model states the probability of a crab having at least one satellite increases by approximately `r ci$coefficients["weight"]`."

### Problem 3.5b

```{r Problem 3.5b}
as.numeric(ci$coefficients["(Intercept)"]+max(hc$weight)*ci$coefficients["weight"])
# We can see the model states the probability of a crab weighing 5.2 kilograms to have more
# than one satellite is approximately 1.533186 (which obviously does not make sense).
cl<-glm(y~weight,data=hc,family=binomial(link=logit))
summary(cl)
eabx<-exp(cl$coefficients["(Intercept)"]+max(hc$weight)*cl$coefficients["weight"])
as.numeric(eabx/(1+eabx))
```

## Problem 3.6

### Problem 3.6a

$P(Y=1)=\frac{e^{3.187-0.5901x}}{1+e^{3.187-0.5901x}}$

Interpretation of *direction*: "As one's political ideology gets more conservative, the probability of one being a Democrat decreases."

### Problem 3.6b, 3.6c, 3.6d

```{r Problem 3.6, message=FALSE}
library(car)
D<-c(5,18,19,25,7,7,2)
n<-D+c(1,3,1,11,10,11,1)
x<-1:7
p<-glm(D/n~x,family=binomial(link=logit),weights=n)
summary(p) # Problem 3.6b
confint(p)["x",]
cat("(",p$coefficients["x"]-qnorm(.975)*summary(p)$coefficients["x","Std. Error"],", ",p$coefficients["x"]+qnorm(.975)*summary(p)$coefficients["x","Std. Error"],")",sep="")
# This confidence interval is smaller than the displayed profile likelihood interval.
as.numeric(p$coefficients["x"]/summary(p)$coefficients["x","Std. Error"])
2*pnorm(as.numeric(p$coefficients["x"]/summary(p)$coefficients["x","Std. Error"]))
# z = -3.772, p = 0.000162                                # Problem 3.6c
# We reject H0 at the alpha = 0.05 level. There is sufficient evidence (p = 0.000162) that
# the effect of x is significant in the model.
Anova(p)   # Problem 3.6d
```

$\chi^2=$ `r Anova(p)[,"LR Chisq"]`, $p=$ `r Anova(p)[,"Pr(>Chisq)"]`

We reject $H_0$ at the $\alpha$ = 0.05 level. There is sufficient evidence (p = 0.0000372) that the effect of x is significant in the model.

### Problem 3.6e

It took the model four (`r summary(p)$iter`) iterations to converge to the maximum likelihood value.

## Problem 3.13

```{r Problem 3.13}
cp<-glm(sat~weight,data=hc,family=poisson)
cp                                         # Problem 3.13a
as.numeric(exp(cp$coefficients["(Intercept)"]+mean(hc$weight)*cp$coefficients["weight"]))
as.numeric(exp(cp$coefficients["weight"])) # Problem 3.13b
# The effect of weight on the model is quite large.
cat("(",cp$coefficients["weight"]-qnorm(.975)*summary(cp)$coefficients["weight","Std. Error"],", ",cp$coefficients["weight"]+qnorm(.975)*summary(cp)$coefficients["weight","Std. Error"],")",sep="")
cat("(",exp(cp$coefficients["weight"]-qnorm(.975)*summary(cp)$coefficients["weight","Std. Error"]),", ",exp(cp$coefficients["weight"]+qnorm(.975)*summary(cp)$coefficients["weight","Std. Error"]),")",sep="")
```

95 percent confidence interval for $\beta$: (`r cp$coefficients["weight"]-qnorm(.975)*summary(cp)$coefficients["weight","Std. Error"]`, `r cp$coefficients["weight"]+qnorm(.975)*summary(cp)$coefficients["weight","Std. Error"]`)

95 percent confidence interval for the multiplicative effect of a 1-kg increase: (`r exp(cp$coefficients["weight"]-qnorm(.975)*summary(cp)$coefficients["weight","Std. Error"])`, `r exp(cp$coefficients["weight"]+qnorm(.975)*summary(cp)$coefficients["weight","Std. Error"])`)